### 2024-09-21 paper/paper.qmd
Initial Commit
template from IC3

---
title: "Marriage Licence Analysis"
subtitle: "STA304 In-class 3"
author: 
  - Amy Jin
thanks: "Code and data are available at: https://github.com/aj3616/STA304IC3"
date: today
date-format: long
abstract: "Data analysis for marrige licence statistics. Graphs and text produced in quarto"
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

# Introduction

This is a analysis about Marriage Licence in different district in Canada, We use @talia, @opendatatorotno, and @rohan.

The remainder of this paper is structured as follows. @sec-mydatasection

# Data {#sec-mydatasection}

Some of our data is of penguins (@fig-marriage), from @palmerpenguins.

```{r}
#| label: fig-marriage
#| fig-cap: Bills of penguins
#| echo: false

data |> 
  ggplot(aes(x = date, y = marriage_licenses)) +
  geom_point()

```

# Discussion

## First discussion point {#sec-first-point}
This is the scatterplot generated by the random sample we selected, this shows an ocsilating trend for marriage licneces
$[0, 500]_{marraigelicence} \land [2010, 2020]_{date}$ and during 2020 and 2025, there is an increasing number of marraige licences and the scatterplots did not clustered in $[0, 500]_{marraigelicence}$.

## Second discussion point
It is safe to assume that the total number of marriages did not vary much, but the average number of licence did increase.

## Third discussion point
It is arguable that less people are getting married in the contemporary time.

## Weaknesses and next steps
We need to perform more statistics analysis to make a more valid conclusion, some next steps can be Times Series analysis, regression estimates and visual representations.

\newpage

\appendix

# Appendix {-}

# Additional data details

\newpage

# References

### 2024-09-21 scripts/00-simulate_data.R

@@ -1,0 +1,40 @@
+#### Preamble ####
+# Purpose: Simulates data
+# Author: Amy Jin
+# Date: 19 September 2024
+# Contact: amyzh.jin@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+
+#### Simulate data ####
+set.seed(304)
+
+# Define the start and end date
+start_date <- as.Date("2018-01-01")
+end_date <- as.Date("2023-12-31")
+
+# Set the number of random dates you want to generate
+number_of_dates <- 100
+
+data <-
+  tibble(
+    dates = as.Date(
+      runif(
+        n = number_of_dates,
+        min = as.numeric(start_date),
+        max = as.numeric(end_date)
+      ),
+      origin = "1970-01-01"
+    ),
+    number_of_marriage = rpois(n = number_of_dates, lambda = 15)
+  )
+
+
+#### Write_csv
+write_csv(data, file = "data/raw_data/simulated.csv")

### 2024-09-21 scripts/01-download_data.R

@@ -1,0 +1,33 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from Open Data Toronto
+# Author: Amy Jin
+# Date: 19 September 2024
+# Contact: amyzh.jin@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+
+
+#### Download data ####
+# get package
+package <- show_package("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
+
+# get all resources for this package
+resources <- list_package_resources("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# load the first datastore resource as a sample
+data <- filter(datastore_resources, row_number()==1) %>% 
+  get_resource()
+
+
+#### Save data ####
+write_csv(data, "data/raw_data/raw_data.csv") 
+

### 2024-09-21 scripts/02-data_cleaning.R

@@ -1,0 +1,26 @@
+#### Preamble ####
+# Purpose: Cleans the raw marriage data into an analysis dataset
+# Author: Amy Jin
+# Date: 19 September 2024
+# Contact: amyzh.jin@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to have downloaded the data
+# Any other information needed? None.
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean data ####
+raw_data <- read_csv("data/raw_data/raw_data.csv")
+
+cleaned_data <-
+  raw_data |>
+  janitor::clean_names() |> 
+  separate(col = time_period,
+            into = c("year", "month"),
+            sep = "-") |> 
+  mutate(date = lubridate::ymd(paste(year, month, "01", sep = "-"))
+         )
+  
+#### Save data ####
+write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-21 scripts/03-test_data.R

@@ -1,0 +1,28 @@
+#### Preamble ####
+# Purpose: Sanity check of the data
+# Author: Amy Jin
+# Date: 19 September 2024
+# Contact: amyzh.jin@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to have simulated data
+# Any other information needed? None.
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+
+#### Test data ####
+data <- read_csv("data/raw_data/simulated.csv")
+
+# Test for negative numbers
+data$number_of_marriage |> min() <= 0
+
+# Test for NAs
+all(is.na(data$number_of_marriage))
+
+
+
+
+
+

### 2024-09-21 scripts/04-model.R

@@ -1,0 +1,37 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Amy Jin
+# Date: 19 September 2024
+# Contact: amyzh.jin@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+
+### Model data ####
+first_model <-
+  stan_glm(
+    formula = flying_time ~ length + width,
+    data = analysis_data,
+    family = gaussian(),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_aux = exponential(rate = 1, autoscale = TRUE),
+    seed = 853
+  )
+
+
+#### Save model ####
+saveRDS(
+  first_model,
+  file = "models/first_model.rds"
+)
+
+

### 2024-09-21 scripts/05-replications.R

@@ -1,0 +1,19 @@
+#### Preamble ####
+# Purpose: Replicated graphs from... [...UPDATE THIS...]
+# Author: Amy Jin
+# Date: 19 September 2024
+# Contact: amyzh.jin@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+# [...UPDATE THIS...]
+
+#### Load data ####
+# [...ADD CODE HERE...]
+
+
+

### 2024-09-21 01-download_data.R
Delete 01-download_data.R
wrong folder

@@ -1,33 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data from Open Data Toronto
-# Author: Amy Jin
-# Date: 19 September 2024
-# Contact: amyzh.jin@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-
-
-#### Download data ####
-# get package
-package <- show_package("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
-
-# get all resources for this package
-resources <- list_package_resources("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
-
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% 
-  get_resource()
-
-
-#### Save data ####
-write_csv(data, "data/raw_data/raw_data.csv") 
-

### 2024-09-21 00-simulate_data.R
simulate data
used student number for seed
changed poison distribution to gamma distribution
got mean from the actual data

@@ -1,0 +1,44 @@
+#### Preamble ####
+# Purpose: Simulates data
+# Author: Amy Jin
+# Date: 19 September 2024
+# Contact: amyzh.jin@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+
+#### Simulate data ####
+set.seed(1007676409)
+
+# Define the start and end date
+start_date <- as.Date("2014-01-01")
+end_date <- as.Date("2024-08-21")
+
+# Set the number of random dates you want to generate
+number_of_dates <- 100
+
+data_ppm = read_csv("data/raw_data/raw_data.csv")
+leadaAmountPpm = as.numeric(data_ppm$`Lead Amount (ppm)`)
+mean_lead <- mean(leadaAmountPpm, na.rm = TRUE)
+
+data <-
+  tibble(
+    dates = as.Date(
+      runif(
+        n = number_of_dates,
+        min = as.numeric(start_date),
+        max = as.numeric(end_date)
+      ),
+      origin = "2014-01-01"
+    ),
+    leadAmountppm = rgamma(n = number_of_dates, shape = 2, rate = 1 / mean_lead)
+    #leadAmountppm = rpois(n = number_of_dates, lambda = mean_lead)
+  )
+
+#### Write_csv
+write_csv(data, file = "data/raw_data/simulated.csv")

### 2024-09-21 00-simulate_data.R
Delete 00-simulate_data.R
wrong folder

@@ -1,44 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates data
-# Author: Amy Jin
-# Date: 19 September 2024
-# Contact: amyzh.jin@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(tidyverse)
-
-
-#### Simulate data ####
-set.seed(1007676409)
-
-# Define the start and end date
-start_date <- as.Date("2014-01-01")
-end_date <- as.Date("2024-08-21")
-
-# Set the number of random dates you want to generate
-number_of_dates <- 100
-
-data_ppm = read_csv("data/raw_data/raw_data.csv")
-leadaAmountPpm = as.numeric(data_ppm$`Lead Amount (ppm)`)
-mean_lead <- mean(leadaAmountPpm, na.rm = TRUE)
-
-data <-
-  tibble(
-    dates = as.Date(
-      runif(
-        n = number_of_dates,
-        min = as.numeric(start_date),
-        max = as.numeric(end_date)
-      ),
-      origin = "2014-01-01"
-    ),
-    leadAmountppm = rgamma(n = number_of_dates, shape = 2, rate = 1 / mean_lead)
-    #leadAmountppm = rpois(n = number_of_dates, lambda = mean_lead)
-  )
-
-#### Write_csv
-write_csv(data, file = "data/raw_data/simulated.csv")

### 2024-09-21 scripts/00-simulate_data.R
simulate data
right folder in scripts

@@ -13,14 +13,18 @@ library(tidyverse)
 
 
 #### Simulate data ####
-set.seed(304)
+set.seed(1007676409)
 
 # Define the start and end date
-start_date <- as.Date("2018-01-01")
-end_date <- as.Date("2023-12-31")
+start_date <- as.Date("2014-01-01")
+end_date <- as.Date("2024-08-21")
 
 # Set the number of random dates you want to generate
 number_of_dates <- 100
+
+data_ppm = read_csv("data/raw_data/raw_data.csv")
+leadaAmountPpm = as.numeric(data_ppm$`Lead Amount (ppm)`)
+mean_lead <- mean(leadaAmountPpm, na.rm = TRUE)
 
 data <-
   tibble(
@@ -30,11 +34,11 @@ data <-
         min = as.numeric(start_date),
         max = as.numeric(end_date)
       ),
-      origin = "1970-01-01"
+      origin = "2014-01-01"
     ),
-    number_of_marriage = rpois(n = number_of_dates, lambda = 15)
+    leadAmountppm = rgamma(n = number_of_dates, shape = 2, rate = 1 / mean_lead)
+    #leadAmountppm = rpois(n = number_of_dates, lambda = mean_lead)
   )
-
 
 #### Write_csv
 write_csv(data, file = "data/raw_data/simulated.csv")

### 2024-09-21 scripts/01-download_data.R
01 download data update

@@ -15,10 +15,10 @@ library(tidyverse)
 
 #### Download data ####
 # get package
-package <- show_package("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
+package <- show_package("8171f7ab-364e-46a7-8a62-08f073ca6d96")
 
 # get all resources for this package
-resources <- list_package_resources("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
+resources <- list_package_resources("8171f7ab-364e-46a7-8a62-08f073ca6d96")
 
 # identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
 datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))

### 2024-09-23 scripts/00-simulate_data.R
update simulate
so that every variable is simulated

@@ -20,14 +20,27 @@ start_date <- as.Date("2014-01-01")
 end_date <- as.Date("2024-08-21")
 
 # Set the number of random dates you want to generate
-number_of_dates <- 100
+n <- 100
 
 data_ppm = read_csv("data/raw_data/raw_data.csv")
 leadaAmountPpm = as.numeric(data_ppm$`Lead Amount (ppm)`)
 mean_lead <- mean(leadaAmountPpm, na.rm = TRUE)
 
+id <- 1:n
+sample_number <- sample(1000000:2000000, n, replace = TRUE)
+sample_date <- sample(seq(start_date, end_date, by = "day"), n, replace = TRUE)
+
+partial_postal_codes <- c("M1L", "M4E", "M4K", "M3B", "M6H", "M4J", "M4M", "M4G", "M6E", "M6P", "M6N", "M2M", "M9B", "M6J")
+partial_postal_code <- sample(partial_postal_codes, n, replace = TRUE)
+
+
 data <-
   tibble(
+    id = id,
+    sample_number = sample_number,
+    sample_date = sample_date,
+    partial_postal_code = partial_postal_code,
+    sample_number = 
     dates = as.Date(
       runif(
         n = number_of_dates,
@@ -37,6 +50,7 @@ data <-
       origin = "2014-01-01"
     ),
     leadAmountppm = rgamma(n = number_of_dates, shape = 2, rate = 1 / mean_lead)
+    
     #leadAmountppm = rpois(n = number_of_dates, lambda = mean_lead)
   )
 

### 2024-09-23 scripts/02-data_cleaning.R

@@ -9,18 +9,23 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(tidyr)
+
 
 #### Clean data ####
 raw_data <- read_csv("data/raw_data/raw_data.csv")
-
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |> 
-  separate(col = time_period,
-            into = c("year", "month"),
-            sep = "-") |> 
-  mutate(date = lubridate::ymd(paste(year, month, "01", sep = "-"))
-         )
+raw_data$`Lead Amount (ppm)` = as.numeric(raw_data$`Lead Amount (ppm)`)
+cleaned_data = na.omit(raw_data)
+cleaned_data = drop_na(cleaned_data)
+#cleaned_data <-
+ # raw_data |>
+  #janitor::clean_names() |> 
+ # separate(col = raw_data$`Sample Date`,
+  #          into = c("year", "month"),
+   #         sep = "-") |> 
+#  mutate(date = lubridate::ymd(paste(year, month, "01", sep = "-"))
+ #        )
   
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")
+

### 2024-09-23 scripts/00-simulate_data.R
clearly comment and structure the script

@@ -13,27 +13,26 @@ library(tidyverse)
 
 
 #### Simulate data ####
+#set seed to my unique student number for reproductibility
 set.seed(1007676409)
+
+# Assume we want 100 simulations, this can be modified upon specification
+n <- 100
 
 # Define the start and end date
 start_date <- as.Date("2014-01-01")
 end_date <- as.Date("2024-08-21")
 
-# Set the number of random dates you want to generate
-n <- 100
-
-data_ppm = read_csv("data/raw_data/raw_data.csv")
-leadaAmountPpm = as.numeric(data_ppm$`Lead Amount (ppm)`)
-mean_lead <- mean(leadaAmountPpm, na.rm = TRUE)
-
+#id is the key for identification just for the data
 id <- 1:n
+#sample number is from the water sample, assume it will be a number between 1000000 and 2000000
 sample_number <- sample(1000000:2000000, n, replace = TRUE)
-sample_date <- sample(seq(start_date, end_date, by = "day"), n, replace = TRUE)
 
+#create a list of the postal codes in toronto and randomly select to simulate
 partial_postal_codes <- c("M1L", "M4E", "M4K", "M3B", "M6H", "M4J", "M4M", "M4G", "M6E", "M6P", "M6N", "M2M", "M9B", "M6J")
 partial_postal_code <- sample(partial_postal_codes, n, replace = TRUE)
 
-
+#join them into a tibble, used the starter code for dates
 data <-
   tibble(
     id = id,
@@ -49,9 +48,9 @@ data <-
       ),
       origin = "2014-01-01"
     ),
+    #used gamma distibution instead of the poison distribution because the mean_lead is a continuous variable,
+    #if we use poisson distibution, all simulations will be 0
     leadAmountppm = rgamma(n = number_of_dates, shape = 2, rate = 1 / mean_lead)
-    
-    #leadAmountppm = rpois(n = number_of_dates, lambda = mean_lead)
   )
 
 #### Write_csv

### 2024-09-23 scripts/00-simulate_data.R
Code styling

@@ -13,7 +13,7 @@ library(tidyverse)
 
 
 #### Simulate data ####
-#set seed to my unique student number for reproductibility
+# set seed to my unique student number for reproductibility
 set.seed(1007676409)
 
 # Assume we want 100 simulations, this can be modified upon specification
@@ -23,23 +23,22 @@ n <- 100
 start_date <- as.Date("2014-01-01")
 end_date <- as.Date("2024-08-21")
 
-#id is the key for identification just for the data
+# id is the key for identification just for the data
 id <- 1:n
-#sample number is from the water sample, assume it will be a number between 1000000 and 2000000
+# sample number is from the water sample, assume it will be a number between 1000000 and 2000000
 sample_number <- sample(1000000:2000000, n, replace = TRUE)
 
-#create a list of the postal codes in toronto and randomly select to simulate
+# create a list of the postal codes in toronto and randomly select to simulate
 partial_postal_codes <- c("M1L", "M4E", "M4K", "M3B", "M6H", "M4J", "M4M", "M4G", "M6E", "M6P", "M6N", "M2M", "M9B", "M6J")
 partial_postal_code <- sample(partial_postal_codes, n, replace = TRUE)
 
-#join them into a tibble, used the starter code for dates
+# join them into a tibble, used the starter code for dates
 data <-
   tibble(
     id = id,
     sample_number = sample_number,
     sample_date = sample_date,
     partial_postal_code = partial_postal_code,
-    sample_number = 
     dates = as.Date(
       runif(
         n = number_of_dates,
@@ -48,8 +47,8 @@ data <-
       ),
       origin = "2014-01-01"
     ),
-    #used gamma distibution instead of the poison distribution because the mean_lead is a continuous variable,
-    #if we use poisson distibution, all simulations will be 0
+    # used gamma distibution instead of the poison distribution because the mean_lead is a continuous variable,
+    # if we use poisson distibution, all simulations will be 0
     leadAmountppm = rgamma(n = number_of_dates, shape = 2, rate = 1 / mean_lead)
   )
 

### 2024-09-23 scripts/01-download_data.R

@@ -21,13 +21,12 @@ package <- show_package("8171f7ab-364e-46a7-8a62-08f073ca6d96")
 resources <- list_package_resources("8171f7ab-364e-46a7-8a62-08f073ca6d96")
 
 # identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+datastore_resources <- filter(resources, tolower(format) %in% c("csv", "geojson"))
 
 # load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% 
+data <- filter(datastore_resources, row_number() == 1) %>%
   get_resource()
 
 
 #### Save data ####
-write_csv(data, "data/raw_data/raw_data.csv") 
-
+write_csv(data, "data/raw_data/raw_data.csv")

### 2024-09-23 scripts/02-data_cleaning.R

@@ -14,18 +14,9 @@ library(tidyr)
 
 #### Clean data ####
 raw_data <- read_csv("data/raw_data/raw_data.csv")
-raw_data$`Lead Amount (ppm)` = as.numeric(raw_data$`Lead Amount (ppm)`)
-cleaned_data = na.omit(raw_data)
-cleaned_data = drop_na(cleaned_data)
-#cleaned_data <-
- # raw_data |>
-  #janitor::clean_names() |> 
- # separate(col = raw_data$`Sample Date`,
-  #          into = c("year", "month"),
-   #         sep = "-") |> 
-#  mutate(date = lubridate::ymd(paste(year, month, "01", sep = "-"))
- #        )
-  
+raw_data$`Lead Amount (ppm)` <- as.numeric(raw_data$`Lead Amount (ppm)`)
+cleaned_data <- na.omit(raw_data)
+cleaned_data <- drop_na(cleaned_data)
+
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")
-

### 2024-09-23 scripts/code_style.R

@@ -1,0 +1,7 @@
+#This script is soley for the purpose of reformatting
+library(styler)
+
+# Reformat an R script
+styler::style_file("scripts/00-simulate_data.R")
+styler::style_file("scripts/01-download_data.R")
+styler::style_file("scripts/02-data_cleaning.R")

### 2024-09-23 scripts/03-test_data.R
perform test to simulated data and analysis data
all checked, safe to proceed with further analysis

@@ -13,14 +13,121 @@ library(tidyverse)
 
 
 #### Test data ####
+#tests developed for simulated data
 data <- read_csv("data/raw_data/simulated.csv")
 
 # Test for negative numbers
-data$number_of_marriage |> min() <= 0
+if(data$`Lead Amount (ppm)` |> min() <= 0){
+  warning("There are negative lead concentration values.")
+}else{
+    message("All lead concentrations are positive.")
+  }
 
 # Test for NAs
-all(is.na(data$number_of_marriage))
+if(all(is.na(data))){
+  warning("There are NA values in the dataset.")
+}else{
+  message("There are no NA in the dataset, safe to proceed.")
+}
 
+# Test for range in SampleNumber
+
+valid_range <- data$`Sample Number` >= 1000000 & data$`Sample Number` <= 30000000
+if (all(valid_range)) {
+  message("All sample numbers are within the valid range.")
+} else {
+  warning("Some sample numbers are outside the valid range.")
+  print(data[!valid_range, ])
+}
+
+# Test for unreasonably high lead concentration
+threshold = 1000
+high_lead <- data$`Lead Amount (ppm)` > threshold
+if (any(high_lead, na.rm = TRUE)) {
+  warning("Some lead concentrations exceed the threshold.")
+  print(data[high_lead, ])
+} else {
+  message("All lead concentrations are within the acceptable range.")
+}
+
+# Test for valid date range
+start_date = "2014-01-01"
+end_date = "2024-09-21"
+valid_dates <- as.Date(data$`Sample Date`) >= as.Date(start_date) & as.Date(data$`Sample Date`) <= as.Date(end_date)
+if (all(valid_dates, na.rm = TRUE)) {
+  message("All dates are within the valid range.")
+} else {
+  warning("Some dates are outside the valid range.")
+  print(data[!valid_dates, ]) 
+}
+
+
+# Test for valid partial postal code
+valid_codes = grepl("^M[0-9]", data$PartialPostalCode)
+if (all(valid_codes)) {
+  message("All partial postal codes are valid.")
+  } else {
+  warning("Some partial postal codes are invalid.")
+  print(data[!valid_codes, ])
+}
+
+# Repeat the tests for actual data
+actual_data <- read_csv("data/analysis_data/analysis_data.csv")
+
+# Test for negative numbers
+if(actual_data$`Lead Amount (ppm)` |> min() <= 0){
+  warning("There are negative lead concentration values.")
[…]

### 2024-09-24 paper/paper.qmd
first draft of paper & readme

@@
 ---
-title: "Marriage Licence Analysis"
-subtitle: "STA304 In-class 3"
+title: "Non-Regulated Residential Lead Testing Program in Toronto"
+subtitle: "Lead Contamination in Water"
 author: 
   - Amy Jin
-thanks: "Code and data are available at: https://github.com/aj3616/STA304IC3"
+thanks: "Code and data are available at: https://github.com/aj3616/Lead-Testing-Program-in-Toronto"
 date: today
 date-format: long
-abstract: "Data analysis for marrige licence statistics. Graphs and text produced in quarto"
+abstract: "The City of Toronto offers a free, non-regulated Residential Lead Testing Program, enabling residents to collect and submit water samples for lead testing.
+Quality of drinking water is an important public health issue.
+This paper examines how time and district affects the lead concentration with respect to safe lead concentration guidelines.
+It also highlights ongoing steps being taken to mitigate lead contamination in the water supply."
 format: pdf
@@
 bibliography: references.bib
+editor: 
+  markdown: 
+    wrap: 72
 ---
@@
 library(tidyverse)
+library(ggplot2)
 
 data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+lead = data$`Lead Amount (ppm)`
 
@@
 
-This is a analysis about Marriage Licence in different district in Canada, We use @talia, @opendatatorotno, and @rohan.
+Lead contamination in drinking water poses significant public health
+risks, particularly for vulnerable populations such as children and
+pregnant women. Even low levels of lead exposure can have serious health
+consequences, including developmental delays and cognitive impairments.
+In response to growing concerns about lead exposure, the City of Toronto
+has implemented a non-regulated Residential Lead Testing Program. This
+initiative allows residents to access free water testing services,
+providing them with the means to assess the lead content in their tap
+water.
+
+The testing process is straightforward: residents can obtain a testing
+kit from any of six Toronto Public Health locations, collect a water
+sample from their home, and return it for analysis. Toronto Water
+processes the samples and provides results directly to the residents.
+Although the City cannot control how the samples are collected, this
+service empowers residents to take action in ensuring the safety of
+their drinking water. This paper will explore the benefits of the
+program, its limitations, and the broader efforts undertaken by the City
+of Toronto to address lead contamination in its water supply system.
+
+We use @talia, @opendatatorotno, and @rohan.
 
@@
 
-Some of our data is of penguins (@fig-marriage), from @palmerpenguins.
+```{r}
+#| label: table1summary
+#| fig-cap: "Table 1: Summary of the lead amount in ppm"
+#| echo: false
+summary(data$`Lead Amount (ppm)`)
+```
 
 ```{r}
-#| label: fig-marriage
-#| fig-cap: Bills of penguins
+#| label: figure1boxplot
+#| fig-cap: "Figure 1: summary of the lead amount in ppm"
 #| echo: false
+boxplot(data$`Lead Amount (ppm)`)
+abline(h = 0.010011423, col = "red")
+```
 
-data |> 
-  ggplot(aes(x = date, y = marriage_licenses)) +
-  geom_point()
+```{r}
+#| echo: false
+threshold <- 0.010011423
+below = data$`Lead Amount (ppm)` < threshold
+sum(below)/10215
+
+```
+
+```{r}
+#| label: figure2ggplot
+#| fig-cap: "Figure 2: Lead Levels changed over time for times seriese analysis"
+#| echo: false
+plot(data$`Sample Date`, data$`Lead Amount (ppm)`, type='l', main="Lead Levels Over Time", xlab="Date", ylab="Lead Amount (ppm)")
+
+```
+
+```{r}
+#| label: figure3ACF
+#| fig-cap: "Figure 3: Autocorrelation function for lead concentrations"
+#| echo: false
+# Autocorrelation function (ACF)
+acf(data$`Lead Amount (ppm)`, na.action = na.omit, main="Autocorrelation of Lead Levels")
+```
+
+```{r}
+#| label: figure4PACF
+#| fig-cap: "Figure 4: Partial Autocorrelation function for lead concentrations"
+#| echo: false
+# Partial Autocorrelation function (PACF)
+pacf(data$`Lead Amount (ppm)`, na.action = na.omit, main="Partial Autocorrelation of Lead Levels")
+
+```
+
+```{r}
+#| label: figure5
+#| fig-cap: "Figure 5: Outliers categorized by district"
+#| echo: false
+boxplot(`Lead Amount (ppm)` ~ PartialPostalCode, data=data, main="Lead Levels by Postal Code", xlab="Partial Postal Code", ylab="Lead Amount (ppm)")
+
+```
+
+```{r}
+#| label: figure6
+#| fig-cap: "Figure 6: Mean lead concentration categorized by first two characters in partial postal code"
+#| echo: false
+# Load necessary library
+library(ggplot2)
+
+# Remove rows with LeadAmountPPM > 4
+filtered_data <- subset(data, `Lead Amount (ppm)` <= 4)
+
+# Categorize by the first two characters of PartialPostalCode
+filtered_data$District <- substr(filtered_data$PartialPostalCode, 1, 2)
+
+# Step 3: Calculate the mean of LeadAmountPPM by District
+mean_lead_by_district <- aggregate(`Lead Amount (ppm)` ~ District, data = filtered_data, FUN = mean)
+print(mean_lead_by_district)
+# Step 4: Construct a bar graph using ggplot2
+ggplot(mean_lead_by_district, aes(x = District, y = `Lead Amount (ppm)`)) +
+  geom_bar(stat = "identity", fill = "steelblue") +
+  labs(title = "Mean Lead Concentration by District",
+       x = "District (First Two Characters of Postal Code)",
+       y = "Mean Lead Amount (ppm)") +
+  theme_minimal()
+
+```
+
+```{r}
+threshold <- 0.010
+high_lead_samples <- data[data$`Lead Amount (ppm)` > threshold, ]
+
+# Count of samples above the threshold
+nrow(high_lead_samples)
+
+# Percentage of samples above threshold
+(nrow(high_lead_samples) / nrow(data)) * 100
 
@@
 
-## First discussion point {#sec-first-point}
-This is the scatterplot generated by the random sample we selected, this shows an ocsilating trend for marriage licneces
-$[0, 500]_{marraigelicence} \land [2010, 2020]_{date}$ and during 2020 and 2025, there is an increasing number of marraige licences and the scatterplots did not clustered in $[0, 500]_{marraigelicence}$.
+## Lead Concentration influencing Health
 
-## Second discussion point
-It is safe to assume that the total number of marriages did not vary much, but the average number of licence did increase.
+-   WHO and Canadian federal guidance on lead concentration in water
 
-## Third discussion point
-It is arguable that less people are getting married in the contemporary time.
+Lead amount The main variable we are examining is the Lead Amount in
+ppm. The distribution of the data is highly right skewed with most
+of its data being very small, as shown in figure 1(@figure1boxplot),
+most data are almost 0 with some outliers which are mostly less than
+2 ppm and 4 of them ranging from 3 to 8.44 which is the maximum
+value as shown in the summary table(@table1summary). We can also
+verify our observation since the summary(@table1summary) shows that
+75% of the data is below 0.001480. Government of Canada had
+specified that the MAC of drinking water is 0.010mg/L(Government
+Canada), which converts to 0.010011423
+part/million(unitconverters.net), this is shown by the red line in
+the figure 1(@figure1boxplot).
+
+-   percentage of data above this threshold
+
+From calculations, there are 97.26% of the data collected indicatin that the lead amount in the water is safe, and it is safe to assume that 75% of the areas in toronto is
+safe of lead in water.
+
+-   plot of the data above this threshold
+
+## How districts affect Lead Concentration
+
+-   graphs of postal code vs lead concentrations
+
+Postal Code Figure 5(@figure5) is a scatterplot of each specific
+partial postal code(first 3 code), we can see that again, most data
+points are around 0, but there are a few outliers with very high
+lead concentration. Figure 5 clarily indicated that which district
+are those outliers in. We can see that the "M4" has two datapoints
+aroun 8 ppm, as well as some between 0 and 2ppm. "M6" also have many
+data between 0 and 2ppm. It is safe to assume that those two district 
+might have a higher overall lead concentration level.
+
+-   extract outliers above the threshold
+
+We will verify the assumption with calculating the mean lead
+concentration when categorized the partial postal codes by only the
+first two numbers. Since outliers might influence the mean, especially
+when in our dataset we are dealing with median 0.000473 and outliers as
+high as 8ppm, we can omit the outliers above 4ppm and perform the data
+analysis.
+
+-   mean value of lead concentration for different districts
+
+Figure 6(@figure6) is the bargraph comparing the mean values
+of lead concentration without outliers above 4ppm and cetegorized by
+only the first two characters of the partial postal code.
+We can see that the top 3 district of the highest lead concentration in
+water is M9, M6 and M5 respectively. This shows that without the data
+above 4ppm,
+
+-   toronto map
+
+## How months and year affect Lead Concentration
+
+-   times series analysis for autocorrelation
+
+Time and Lead In this dataset, we have variables time and lead amount, therefore, we can perform times series analysis(STA457).
+As shown in figure 2(@figure2ggplot), we can see that except for some sharp increase in 2014, 2019, 2021, 2023, the overall trend is stays
+around 0 and remains positive.
+
+From the ACF and PACF figures, we can conclude that the previous values
+of lead concentrations does not impact future values. For ACF, when lag
+is 0, ACF is always 1 but for all other lags in ACF(@figure3ACF) and
+PACF(@figure4PACF) they are all within the dotted blue lines which is
+the evidence that there are no significant autocorrelation nor
+autoregressive realtionship present in the data.
+
+This observation is reasonable because the samples are collected in
+different district in Toronto, and it is safe to assume that one
+incident of a very high lead concentration might again only be outliers
+instead of some factors that will effect the overall water quality in
+Toronto.
+
+-   graphs of months vs mean lead concentration
+
+-   graphs of years vs mean lead concentration
+
+-   extract outliers above the threshold
 
 ## Weaknesses and next steps
-We need to perform more statistics analysis to make a more valid conclusion, some next steps can be Times Series analysis, regression estimates and visual representations.
+
+-   the data is from many different water source, they are not
+    necessarily direct drinking water since how the water sample
+    collected was not controled. However, the guides are for drinking
+    water.
+
+-   importance of pre-treatment such as filter
 
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 

### 2024-09-24 scripts/03-test_data.R

@@ -13,20 +13,20 @@ library(tidyverse)
 
 
 #### Test data ####
-#tests developed for simulated data
+# tests developed for simulated data
 data <- read_csv("data/raw_data/simulated.csv")
 
 # Test for negative numbers
-if(data$`Lead Amount (ppm)` |> min() <= 0){
+if (data$`Lead Amount (ppm)` |> min() <= 0) {
   warning("There are negative lead concentration values.")
-}else{
-    message("All lead concentrations are positive.")
-  }
+} else {
+  message("All lead concentrations are positive.")
+}
 
 # Test for NAs
-if(all(is.na(data))){
+if (all(is.na(data))) {
   warning("There are NA values in the dataset.")
-}else{
+} else {
   message("There are no NA in the dataset, safe to proceed.")
 }
 
@@ -41,7 +41,7 @@ if (all(valid_range)) {
 }
 
 # Test for unreasonably high lead concentration
-threshold = 1000
+threshold <- 1000
 high_lead <- data$`Lead Amount (ppm)` > threshold
 if (any(high_lead, na.rm = TRUE)) {
   warning("Some lead concentrations exceed the threshold.")
@@ -51,22 +51,22 @@ if (any(high_lead, na.rm = TRUE)) {
 }
 
 # Test for valid date range
-start_date = "2014-01-01"
-end_date = "2024-09-21"
+start_date <- "2014-01-01"
+end_date <- "2024-09-21"
 valid_dates <- as.Date(data$`Sample Date`) >= as.Date(start_date) & as.Date(data$`Sample Date`) <= as.Date(end_date)
 if (all(valid_dates, na.rm = TRUE)) {
   message("All dates are within the valid range.")
 } else {
   warning("Some dates are outside the valid range.")
-  print(data[!valid_dates, ]) 
+  print(data[!valid_dates, ])
 }
 
 
 # Test for valid partial postal code
-valid_codes = grepl("^M[0-9]", data$PartialPostalCode)
+valid_codes <- grepl("^M[0-9]", data$PartialPostalCode)
 if (all(valid_codes)) {
   message("All partial postal codes are valid.")
-  } else {
+} else {
   warning("Some partial postal codes are invalid.")
   print(data[!valid_codes, ])
 }
@@ -75,16 +75,16 @@ if (all(valid_codes)) {
 actual_data <- read_csv("data/analysis_data/analysis_data.csv")
 
 # Test for negative numbers
[…]

### 2024-09-24 scripts/code_style.R

@@ -5,3 +5,4 @@ library(styler)
 styler::style_file("scripts/00-simulate_data.R")
 styler::style_file("scripts/01-download_data.R")
 styler::style_file("scripts/02-data_cleaning.R")
+styler::style_file("scripts/03-test_data.R")

### 2024-09-27 paper/paper.qmd
Written paper qmd and update citations

--- 2024-09-21 paper/paper.qmd
@@
 ---
-title: "Marriage Licence Analysis"
-subtitle: "STA304 In-class 3"
+title: "Lead Contamination Testing: Is water in Toronto safe for drinking?"
+subtitle: "Water in Some Districts is more Contaminated than others"
 author: 
   - Amy Jin
-thanks: "Code and data are available at: https://github.com/aj3616/STA304IC3"
+thanks: "Code and data are available at: https://github.com/aj3616/Lead-Testing-Program-in-Toronto"
 date: today
 date-format: long
-abstract: "Data analysis for marrige licence statistics. Graphs and text produced in quarto"
+abstract: "The City of Toronto offers a free, non-regulated Residential Lead Testing Program, where residents collect and submit water samples.
+This paper examines how time and district affects the lead concentration with respect to safe lead concentration guidelines.
+Quality of drinking water is an important public health issue, since too much abosorbtion of lead might have detrimental effects especially for young children.
+It was found that the lead concentration does not have noticeble seasonality or autocorrelation, some areas in toronto tends to have higher lead concentrations than the rest."
 format: pdf
@@
 library(tidyverse)
+library(ggplot2)
+library(knitr)
+library(gt)
+library(lubridate)  # For handling date extraction
+library(dplyr)
 
 data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+lead = data$`Lead Amount (ppm)`
+district = data$PartialPostalCode
+
+blueishgrey <- rgb(96, 139, 182, maxColorValue = 255)
 
@@
 
-This is a analysis about Marriage Licence in different district in Canada, We use @talia, @opendatatorotno, and @rohan.
+Lead contamination in drinking water poses significant public health
+risks, particularly for vulnerable populations such as children and
+pregnant women(Government Canada).
+
+This paper takes in raw data from the City of Toronto(Open Data
+Toronto), where they implemented a non-regulated program called City's
+Residential Lead Testing Program. More information on this program can
+be explored at: www.toronto.ca/leadpipes. The results respond to the
+concerns about lead exposure in the city of Toronto.
+
+While significant research has extensively documented the public health
+impacts of lead exposure, particularly focusing on its toxic effects on
+vulnerable populations like children and pregnant women (Levallois et
+al., 2018)(Jarvis & Fawell, 2021), a critical gap remains in
+understanding how lead concentrations in urban drinking water systems
+fluctuate both spatially and temporally.
+
+This paper pre-processed the raw data and the cleaned data was used for
+visual representations and statistical analysis includes outliers
+analysis, time series analysis and linear regression analysis.
+
+The results from the analysis show that there is no clear seasonality,
+which refers to recurring patterns or trends in data that repeat over
+specific time intervals, and no significant autocorrelation, which
+indicates a lack of correlation between current and past lead
+concentrations. In other words, future lead concentrations do not depend
+on previous lead levels, and there is insufficient evidence to suggest
+that certain months consistently exhibit higher trends of lead
+concentrations compared to others.
+
+However, some district does have higher trend of lead concentration.
+There are clusters of outliers within some specific district and some
+district have higher mean lead concentrations than other. It is safe to
+say that geographic locations and community water maintenance have an
+effect on water safety.
+
+This topic is important because even low levels of lead exposure can
+have serious health consequences, including developmental delays,
+cognitive impairments and even detrimental effects(Government Canada).
+
+We use @talia for data analysis, @opendatatorotno for collected data
+set, and @rohan for directions and template.
 
 The remainder of this paper is structured as follows. @sec-mydatasection
+Data, further explains the data set we are working with and the
+prepossessing done for further analysis. A discussion with the
+measurement used when collecting the data is also included. Graphs and
+tables were used to show important features of the data set.
+@sec-mydiscussion Discussion, talks about, firstly, to what extent will
+Lead Concentration influence Health and how much concentration will be
+categorized as “dangerous”. Secondly, the paper went into a detailed
+analysis of how districts affect lead concentration. Graphs and outliers
+analysis was used to present the result. Thirdly, the paper performs
+Times Series analysis to analyze the lead concentration over time.
+Lastly, there is a subsection discussing weakness of the paper and some
+further steps to further support the question.
 
@@
 
-Some of our data is of penguins (@fig-marriage), from @palmerpenguins.
+## Tools Used
 
+#TODO: change library citation R (R Core Team 2023) was used for the
+data analysis and the productions of all graphs and tables. The
+libraries tidyverse (Wickham et al. 2019), here (J. B. Kirill Müller
+2020), dplyr (Hadley Wickham 2023), tibble (R. F. Kirill Müller Hadley
+Wickham 2023), janitor (Sam Firke 2023), ggplot2 (Wickham 2016), and
+knitr (Xie 2023) were also used for organizing data and performing
+analysis.
+
+## Data Set and Context
+The data set "Non Regulated Lead Sample" by Open Data Toronto is to
+address the concerns Toronto residents have for their drinking water.
+The data set contains information on lead concentrations in water
+samples. Each row is from a water sample, with corresponding information
+on the date, partial postal code where the sample was provided, and the
+main variable of interest, lead concentration in parts per million
+(ppm).
+
+Lead exposure is an important public health issue. High levels of lead
+in drinking water can pose severe health risks, particularly to
+vulnerable populations.(WHO) Chronic exposure to lead can result in
+developmental delays, neurological damage, and a range of other health
+problems.(WHO) Tracking lead concentrations in water is crucial for
+ensuring safe water quality and identifying areas requiring intervention
+to effectively manage health risks.
+
+There are similar similar data sets that addresses the problem of hard
+water contamination. One example is in state of Michigan, USA, MIDHHS
+collected similar sample with months, years and lead
+concentrations(MIDHHS). There are also other data set that analyze a
+different metal in water. However, this data set is used because lead is
+a very common metal in hard water with severe damage to public health
+when overdose, and being a resident in city of Toronto, this should be a
+concern that needs attention.
+
+## Variables
+
+There are five variables in total: \_id, Sample Number, Sample Date,
+PartialPostalCode and Lead Amount. While the first two are keys of this
+data set for identification purpose, the rest are our variable of
+interest:
+
+**Date**: Represents the date on which the water sample was collected,
+can be used to detect trends and seasonal variations.
+
+**Partial Postal Code:** A truncated form of the postal code indicating
+the general area where the water sample was taken, can be used to
+identify potential geographical clusters.
+
+**Lead Concentration (ppm):** The measured amount of lead in the water
+sample, expressed in parts per million. This is the primary variable of
+interest, as it directly relates to public health risks associated with
+lead exposure. We can also further analyze combinations of lead
+concentration with other variables.
+
+Other than the given variables, there are three additional variables
+constructed during analysis: Months: Extracted from the Date variable to
+analyze seasonal patterns in lead concentration. Years: Extracted from
+the Date variable to evaluate trends over time and aggregate the data
+for yearly analysis. District (First Two Digits of Postal Code): Created
+from the Partial Postal Code to group data into broader geographic areas
+within the city, aiding in the spatial analysis of lead concentration.
+
+## Cleaning
+
+For the purpose of data analysis the raw data set went through some
+pre-processing: NA Values: All missing (NA) values were removed to
+ensure accurate analysis. Lead Concentration: The Lead Amount (ppm)
+variable was cleaned to remove any anomalies and ensure that values were
+properly numeric. Date Variable Type: The Date variable was converted to
+Date type in R to facilitate temporal analysis and aggregation by months
+and years.
+
+After cleaning the data looks like:
+
+**Date(Date):** A Date type, ranges from 2014-01-01 to 2024-08-21,
+covering over a decade of water sampling data.
+
+**District(chr):** A Character type, represents various regions within
+the city of Toronto based on the first two characters of the Partial
+Postal Code.
+
+**Lead Concentration(num):** A num type, which represents the summary
+statistics of lead concentration typically include measures like the
+minimum, maximum, mean, median, and standard deviation of lead levels
+across the data set.
+
+## Summary
+Table 1 is a summary table for num type variable lead amount.
+We can see that the minimum and maximum value differs by 168,800 times, which is very significant.
+Therefore, we can assume that there are water samples that are very clean in terms of lead concentration and there are samples that are contaminated.
+We can also see that the mean and median differs a lot.
+Which means that the distribution of this data set is highly right skewed.
+It also shows evidence that there are many outliers that will elevated mean lead concentration level in this data set.
+
+```{r, echo=FALSE}
+#| label: tbl-summary
+#| tbl-cap: "Distribution Statistics Measurements of Lead Amount"
+#| echo: false
+# Example summary statistics
+lead_summary <- summary(data$`Lead Amount (ppm)`)
+
+# Convert summary to a data frame
+summary_df <- data.frame(
+  Metric = names(lead_summary),
+  Value = as.numeric(lead_summary)
+)
+
+# Create a table using gt
+gt_table <- summary_df %>%
+  gt() %>%
+  tab_header(
+    title = "Summary of Lead Amount (ppm)"
+  ) %>%
+  cols_label(
+    Metric = "Statistic",
+    Value = "Value"
+  ) %>%
+  fmt_number(
+    columns = "Value",
+    decimals = 6
+  )
+
+# Print the gt table
+gt_table
+
+```
+## Relationships
+These relationships are upon further discussion in @sec-mydiscussion.
+
+**Lead Concentration and Time (Months/Years)**: Analyzing lead levels
+over time can reveal temporal trends, such as reductions in lead levels
+following interventions, seasonal spikes, or gradual increases over
+certain periods.
+
+**Lead Concentration and Location (District)**: By examining lead
+concentrations across districts, one can identify areas with
+consistently high or low lead levels, indicating possible regional
+differences in water quality or pipe infrastructure.
+
+**Seasonality**: The monthly variable can help detect seasonal patterns,
+such as increased lead concentration during warmer months due to factors
+like water temperature affecting pipe corrosion.
+
+**Temporal Trends**: By grouping data by year, any long-term trends,
+such as a decrease in lead levels due to policy interventions, can be
+detected.
+
+## Measurement
+In this dataset, we are interested in the concentration of lead in drinking water across different regions of a city and over time.
+The measurement process transforms a real-world phenomenon—lead contamination in water—into quantitative entries in a dataset that can be analyzed for trends.
+
+**Lead concentration:** Measured in parts per million (ppm), indicating the amount of lead in the water relative to its volume.
+This choice of unit is small enough to capture low but potentially harmful concentrations and is standard in water quality testing.
+This transformation from the chemical presence of lead to a numeric value is a key aspect of measurement.
+
+**Sample Date:** The date the sample was taken is recorded to allow for analysis over time, detecting patterns.
+
+**Partial Postal Code:** Postal code represents where the sample was collected, allowing comparisons between different geographic areas.
+This made it easy to get a result using data analysis to public health control organizations.
+
+Some challenges are that this is a non-regulated program, therefore the measurement process is not controlled, decreasing accuracy and precision.
+However, this is minimized by providing a detailed instructions with the sample kit used to collect the water samples and the deviation will only affect Lead concentration.
+Date and district are accurate.
+
+Choices about measurement such as the instruments used, units of measurement, and protocols for sample collection—affect how well the data set represents the real-world phenomenon of lead contamination.
+
+# Discussion {#sec-mydiscussion}
+
+## Lead Concentration influencing Health
+
+### WHO and Canadian federal guidance on lead concentration in water
+
+The main variable we are examining is the Lead Amount in
+ppm. The distribution of the data is highly right skewed with most of
+its data being very small, as shown in (@fig-boxplot), most
+data are almost 0 with outliers which are mostly less than 2 ppm
+and 4 of them ranging from 3 to 8.44 which is the maximum value as shown
+in the summary Table 1. We can also verify our observation
+since the summary Table 1 shows that 75% of the data is below
+0.001480. Government of Canada had specified that the maximum acceptable concentration (MAC) of drinking
+water is 0.010mg/L(Government Canada), which converts to 0.010011423
+part/million(unitconverters.net), this is shown by the red line in the
+figure 1(@fig-boxplot).
 ```{r}
-#| label: fig-marriage
-#| fig-cap: Bills of penguins
+#| label: fig-boxplot
+#| fig-cap: Boxplot showing the lead concentration above the saftey threshold
 #| echo: false
 
-data |> 
-  ggplot(aes(x = date, y = marriage_licenses)) +
-  geom_point()
+# Set the threshold value for the horizontal line
+threshold_value <- 0.010011423
+
+# Create the boxplot using ggplot2
+ggplot(data, aes(y = `Lead Amount (ppm)`)) +
+  geom_boxplot(fill = "skyblue") +
+  geom_hline(yintercept = threshold_value, color = "red", linetype = "dashed") +
+  labs(title = "Boxplot of Lead Amount (ppm) with Threshold Line",
+       y = "Lead Amount (ppm)") +
+  theme_minimal()
 
@@
 
-# Discussion
+### Percentage of data above this threshold
 
-## First discussion point {#sec-first-point}
-This is the scatterplot generated by the random sample we selected, this shows an ocsilating trend for marriage licneces
-$[0, 500]_{marraigelicence} \land [2010, 2020]_{date}$ and during 2020 and 2025, there is an increasing number of marraige licences and the scatterplots did not clustered in $[0, 500]_{marraigelicence}$.
+From calculations by R, there are 97.26% of the data collected indicating that
+the lead amount in the water is safe.
+As shown in (@tbl-threshold), there are only 280 counts of water sample collected that is above MAC, which takes up only 2.74% of the whole data set.
+Even though this seems like a good result, those 280 measurements still effects people and their family's health since those lead concentrations will have negative effect on health(WHO)(Canadian Government).
+```{r, echo=FALSE}
+#| label: tbl-threshold
+#| tbl-cap: "Number and Percentage of data above MAC"
+#| echo: false
+# Define the threshold
+threshold <- 0.010
 
-## Second discussion point
-It is safe to assume that the total number of marriages did not vary much, but the average number of licence did increase.
+# Subset of samples above the threshold
+high_lead_samples <- data[data$`Lead Amount (ppm)` > threshold, ]
 
-## Third discussion point
-It is arguable that less people are getting married in the contemporary time.
+# Calculate the count and percentage
+count_above_threshold <- nrow(high_lead_samples)
+percentage_above_threshold <- (count_above_threshold / nrow(data)) * 100
 
-## Weaknesses and next steps
-We need to perform more statistics analysis to make a more valid conclusion, some next steps can be Times Series analysis, regression estimates and visual representations.
+# Create a summary table
+summary_table <- data.frame(
+  Metric = c("Count of Samples Above Threshold", "Percentage of Samples Above Threshold"),
+  Value = c(count_above_threshold, round(percentage_above_threshold, 2))
+)
+
+# Use gt to create a formatted table
+summary_table %>%
+  gt() %>%
+  tab_header(
+    title = "Summary of Lead Concentration Above Threshold"
+  ) %>%
+  fmt_number(
+    columns = "Value",
+    decimals = 2
+  )
+
+```
+
+### plot of the data above this threshold
+To visualize those data above MAC, data below MAC and above 2 were filtered out, because from (@fig-boxplot), we can see that there are only 4 data points above 2, therefore, with that in mind it is safe to temporarily ignore to get a scale that will better show the most data points above MAC.
+A scatterplot was used with x axis being date and y axis being lead concentration give a graph that spreads out lead concentrations instead of on a 1-dimensional line.
+From (@fig-between) we can see that there are clusters of data points below 0.25 and for higher lead concentrations, there are less data points and more spread out than those below 0.25.
+
+```{r}
+#| label: fig-between
+#| fig-cap: "Data between 2 and MAC representing with Date"
+#| echo: false
+# Define the threshold value
+threshold <- 0.010011423
+upper_threshold = 2
+
+# Filter data for values above the threshold
+data_above_threshold <- data %>%
+  filter(upper_threshold >`Lead Amount (ppm)` & `Lead Amount (ppm)` > threshold)
+
+# Plot the data using ggplot2 (scatter plot example)
+ggplot(data_above_threshold, aes(x = `Sample Date`, y = `Lead Amount (ppm)`)) +
+  geom_point(color = "red") +  # Use red points to highlight threshold exceedance
+  labs(title = "Lead Levels Above Threshold",
+       x = "Sample Date",
+       y = "Lead Amount (ppm)") +
+  theme_minimal()
+
+```
+
+## How districts affect Lead Concentration
+
+### graphs of postal code vs lead concentrations
+
+Nine boxplot (@fig-boxplotdistrict) each representing a district, which is the constructed variable from the first two letters of partial postal code.
+Again, most data points are around 0, and the outliers streched the scale so that this way we can easily analyze the outliers.
+The districts containing outliers were also shown in (@fig-boxplotdistrict).
+We can see that the "M4" has two data points above 8 ppm, as
+well as a cluster between 0 and 1 ppm. "M6" also have many data between 0 and
+2ppm. It is safe to assume that these two districts have water sources with high lead concentrations.
+```{r}
+#| label: fig-boxplotdistrict
+#| fig-cap: "Boxplot Showing Outliers categorized by district"
+#| echo: false
+
+# Create a new data frame with District extracted
+data_district <- data.frame(
+  District = substr(data$PartialPostalCode, 1, 2),
+ lead = data$`Lead Amount (ppm)`
+)
+
+# Create the boxplot using ggplot2 without modifying the original data
+ggplot(data_district, aes(x = District, y = lead)) +
+  geom_boxplot(fill = "black", color = blueishgrey, outlier.color = "red", outlier.shape = 16) +
+  labs(title = "Lead Levels by District",
+       x = "District (First Two Characters of Postal Code)",
+       y = "Lead Amount (ppm)") +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  # Center and bold title
+    axis.title.x = element_text(size = 14, face = "bold"),  # Bold x-axis title
+    axis.title.y = element_text(size = 14, face = "bold"),  # Bold y-axis title
+    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),  # Rotate x-axis labels
+    axis.text.y = element_text(size = 12)
+  )
+
+```
+
+### extract outliers above the threshold
+
+We will verify which district have a higher overall lead concentration by calculating the mean lead concentration when still categorize the partial postal codes by only the
+first two numbers. Since outliers might influence the mean, especially
+when in our dataset we are dealing with median 0.000473 and outliers as
+high as 8ppm, the outliers above 2 ppm were omitted before caculating the mean lead concentration.
+
+### mean value of lead concentration for different districts
+
+The bargraph (@fig-meandistrict) compares the mean values of lead
+concentration without outliers above 2 ppm and categorized by only the
+first two characters of the partial postal code. We can see that the top
+two district of the highest lead concentration in water is M9 and M6 respectively.
+This shows that without the data above 2 ppm, M4 does not have a overall high lead concentration, but only some very high outliers.
+However, M9 and M6 will be the two most dangerous district in Toronto in terms of lead concentration in water.
+Even though all the mean values shown in (@tbl-meandistrict) are all below MAC by Canadian Government, the M9 and M6 district might need more attention in the topic of water saftey given that the rest districts perform better especially M8.
+
+```{r}
+#| label: tbl-meandistrict
+#| tbl-cap: "Specific Numbers of the mean for each district"
+#| echo: false
+
+# Remove rows with LeadAmountPPM > 4
+filtered_data <- subset(data, `Lead Amount (ppm)` <= 2)
+
+# Categorize by the first two characters of PartialPostalCode
+filtered_data$District <- substr(filtered_data$PartialPostalCode, 1, 2)
+
+# Step 3: Calculate the mean of LeadAmountPPM by District
+mean_lead_by_district <- aggregate(`Lead Amount (ppm)` ~ District, data = filtered_data, FUN = mean)
+
+# Create the table using gt
+mean_lead_by_district %>%
+  gt() %>%
+  tab_header(
+    title = "Mean Lead Concentration by District"
+  ) %>%
+  cols_label(
+    District = "District",
+    `Lead Amount (ppm)` = "Mean Lead Amount (ppm)"
+  ) %>%
+  fmt_number(
+    columns = `Lead Amount (ppm)`,
+    decimals = 5
+  )
+
+```
+
+```{r}
+#| label: fig-meandistrict
+#| fig-cap: "Mean lead concentration categorized by first two characters in partial postal code"
+#| echo: false
+# Step 4: Construct a bar graph using ggplot2
+ggplot(mean_lead_by_district, aes(x = District, y = `Lead Amount (ppm)`)) +
+  geom_bar(stat = "identity", fill = blueishgrey) +
+  labs(title = "Mean Lead Concentration by District",
+       x = "District (First Two Characters of Postal Code)",
+       y = "Mean Lead Amount (ppm)") +
+  theme_minimal()
+```
+
+## How months and year affect Lead Concentration
+
+### Times Series analysis for autocorrelation{#sec-timeseries}
+
+In this dataset, we have variables time and lead amount,
+therefore, we can perform times series analysis. As shown in
+(@fig-timeseries), we can see that except for some sharp increase
+in 2014, 2019, 2021, 2023, the overall trend is stays around 0 and
+remains positive. There are no increasing or decreasing trend overtime and no obvious seasonality present.
+
+From the ACF and PACF figures, we can conclude that the previous values
+of lead concentrations does not impact future values. For ACF, when lag
+is 0, ACF is always 1 but for all other lags in ACF(@fig-ACF) and
+PACF(@fig-PACF) they are all within the dotted blue lines which is
+the evidence that there are no significant correlation at any lags present in the data.
+
+This observation is reasonable because the samples are collected in
+different district in Toronto, and it is safe to assume that one
+incident of a very high lead concentration might again only be outliers
+instead of due some factors that will effect the overall water quality in
+Toronto.
+
+Lead Concentration in water can vary overtime(WHO), therefore the high lead concentration outliers might not be caused by correlation instead, "a probability-based adaptive sampling plan should be used to access exposure(WHO)".
+
+```{r}
+#| label: fig-timeseries
+#| fig-cap: "Lead Levels changed over time for times seriese analysis"
+#| echo: false
+
+# Create a line plot using ggplot2
+ggplot(data, aes(x = `Sample Date`, y = `Lead Amount (ppm)`)) +
+  geom_line(color = blueishgrey, linewidth = 0.5) +
+  labs(title = "Lead Levels Over Time",
+       x = "Date",
+       y = "Lead Amount (ppm)") +
+  theme_minimal()
+
+```
+
+```{r}
+#| label: fig-ACF
+#| fig-cap: "Autocorrelation function for lead concentrations"
+#| echo: false
+# Autocorrelation function (ACF)
+# Calculate ACF (autocorrelation function)
+acf_result <- acf(data$`Lead Amount (ppm)`, na.action = na.omit, plot = FALSE)
+
+# Extract the lags and acf values from the result
+acf_data <- data.frame(
+  Lag = acf_result$lag,
+  ACF = acf_result$acf
+)
+
+# Create the ACF plot using ggplot2
+ggplot(acf_data, aes(x = Lag, y = ACF)) +
+  geom_bar(stat = "identity", fill = "white", color = "black") +
+  geom_hline(yintercept = c(0.2, -0.2), linetype = "dashed", color = "blue") +  # Add reference lines
+  labs(title = "Autocorrelation of Lead Levels",
+       x = "Lag",
+       y = "Autocorrelation") +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),  # Center and bold title
+    axis.title = element_text(size = 14),
+    axis.text = element_text(size = 12)
+  )
+
+```
+
+```{r}
+#| label: fig-PACF
+#| fig-cap: "Partial Autocorrelation function for lead concentrations"
+#| echo: false
+# Partial Autocorrelation function (PACF)
+pacf = pacf(data$`Lead Amount (ppm)`, na.action = na.omit, main="Partial Autocorrelation of Lead Levels")
+
+```
+
+### graphs of months vs mean lead concentration
+With all data points, we do not see obvious trends over time.
+Here we examine closely for the constructed variable month from Date.
+(@fig-month) is the boxplot showing the outliers, but there are not enough evidence to show that there are specific month that will have a higher lead concentration.
+Therefore, we perform a similar filtering on data that is larger than 2 ppm.
+```{r}
+#| label: fig-month
+#| fig-cap: "Boxplots that shows the outliers categorized by Months"
+#| echo: false
+# Extract month as a factor for visualization
+data_by_month <- data %>%
+  mutate(Month = month(`Sample Date`, label = TRUE, abbr = TRUE))
+
+# Create a boxplot of lead concentration by month with red outliers
+ggplot(data_by_month, aes(x = Month, y = `Lead Amount (ppm)`)) +
+  geom_boxplot(outlier.color = "red", outlier.shape = 16, fill = blueishgrey) +
+  labs(title = "Lead Concentration by Month",
+       x = "Month",
+       y = "Lead Amount (ppm)") +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+    axis.text.x = element_text(size = 12),
+    axis.text.y = element_text(size = 12)
+  )
+
+```
+
+Bar graph (@fig-meanmonth) shows the mean lead concentration for each month in the year.
+April and September has higher lead concentrations that is about 500% of the other months.
+However, this graph does not provide enough evidence to prove seasonality.
+We need we examine whether April and September for every year will repeatedly have higher measured lead concentration.
+This will be further discussed in @sec-future.
+
+```{r}
+#| label: fig-meanmonth
+#| fig-cap: "Mean Lead Concentration Without Data Greater than 2 ppm categorized by Months"
+#| echo: false
+# Extract month and calculate mean lead concentration by month
+mean_lead_by_month <- data %>%
+  mutate(Month = month(`Sample Date`, label = TRUE, abbr = TRUE)) %>%
+  group_by(Month) %>%
+  summarise(mean_lead = mean(`Lead Amount (ppm)`))
+
+# Create the plot of mean lead concentration by month
+ggplot(mean_lead_by_month, aes(x = Month, y = mean_lead)) +
+  geom_bar(stat = "identity", fill = blueishgrey) +
+  labs(title = "Mean Lead Concentration by Month",
+       x = "Month",
+       y = "Mean Lead Amount (ppm)") +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+    axis.text.x = element_text(size = 12),
+    axis.text.y = element_text(size = 12)
+  )
+
+```
+
+### graphs of years vs mean lead concentration
+Similar to months, we also constructed a variable call year.
+However, both the boxplots and the mean values without data below 2 ppm does not show trend from year to year.
+This aligns with our conclusion in @sec-timeseries.
+```{r}
+#| label: fig-year
+#| fig-cap: "Boxplots that shows the outliers categorized by Year"
+#| echo: false
+# Create a new data frame with the extracted year
+data_by_year <- data.frame(
+  Year = year(data$`Sample Date`),
+  lead = data$`Lead Amount (ppm)`
+)
+
+# Create a ggplot with lead concentration categorized by years
+ggplot(data_by_year, aes(x = factor(Year), y = lead)) +
+  geom_boxplot(fill = "skyblue", color = blueishgrey,outlier.color = "red") +
+  labs(title = "Lead Concentration by Year",
+       x = "Year",
+       y = "Lead Amount (ppm)") +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+    axis.text.x = element_text(size = 12),
+    axis.text.y = element_text(size = 12)
+  )
+
+```
+```{r}
+#| label: fig-meanyear
+#| fig-cap: "Mean Lead Concentration Without Data Greater than 2 ppm categorized by Year"
+#| echo: false
+# Filter data with threshold of 2 ppm
+filtered_data <- data %>%
+  filter(`Lead Amount (ppm)` <= 2)
+
+# Extract year and calculate mean lead concentration by year
+mean_lead_by_year <- filtered_data %>%
+  mutate(Year = year(`Sample Date`)) %>%
+  group_by(Year) %>%
+  summarise(mean_lead = mean(`Lead Amount (ppm)`))
+
+# Create the plot of mean lead concentration by year
+ggplot(mean_lead_by_year, aes(x = factor(Year), y = mean_lead)) +
+  geom_bar(stat = "identity", fill = blueishgrey, color = blueishgrey) +
+  labs(title = "Mean Lead Concentration by Year",
+       x = "Year",
+       y = "Mean Lead Amount (ppm)") +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
+    axis.text.x = element_text(size = 12),
+    axis.text.y = element_text(size = 12)
+  )
+
+```
+
+## Weaknesses and next steps{#sec-future}
+
+A key weakness of the current dataset is that it originates from a variety of water sources and is not necessarily limited to direct drinking water.
+Since the water sample collection process was not controlled, it is uncertain whether all samples reflect tap water quality.
+Despite this, the guidelines used for analysis are intended for drinking water safety, which introduces a potential mismatch between the data and its intended use.
+The program is non-regulated, leaving the sampling process up to residents and without any standardization; as a result, the reliability and comparability of the data across different samples may be compromised.
+
+Considering these weaknesses, it is important to recognize the role of pre-treatment, such as water filtration systems, in affecting lead concentration.
+This necessitates the documentation of whether samples are taken before or after filtration, which could significantly influence the interpretation of lead levels in relation to health guidelines.
+To improve the reliability of future data collection, a probability-based adaptive sampling plan should be considered.
+This approach would ensure more representative sampling of tap water from various districts, providing a better understanding of the exposure to lead in drinking water.
+Additionally, further analysis on the seasonality of lead levels could reveal patterns influenced by environmental factors such as temperature, water flow rates, and seasonal water treatment processes.
+Identifying these trends is crucial for developing targeted interventions to reduce lead exposure throughout the year.
 
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
+The color used in graphs :
+blueishgrey <- rgb(96, 139, 182, maxColorValue = 255) is the color of lead in reality.
 

### 2024-09-27 paper/paper.qmd
Update README and bibliography
removed first person words and improved grammar

@@
 ---
-title: "Lead Contamination Testing: Is water in Toronto safe for drinking?"
-subtitle: "Water in Some Districts is more Contaminated than others"
+title: "Lead Contamination Testing: Is Water in Toronto Safe for Drinking?"
+subtitle: "Water in Some Districts is More Contaminated Than Others"
 author: 
@@
 abstract: "The City of Toronto offers a free, non-regulated Residential Lead Testing Program, where residents collect and submit water samples.
+Quality of drinking water is an important public health issue, since too much abosorbtion of lead might have detrimental effects especially for young children.
 This paper examines how time and district affects the lead concentration with respect to safe lead concentration guidelines.
-Quality of drinking water is an important public health issue, since too much abosorbtion of lead might have detrimental effects especially for young children.
-It was found that the lead concentration does not have noticeble seasonality or autocorrelation, some areas in toronto tends to have higher lead concentrations than the rest."
+It was found that the lead concentration does not have noticeble seasonality or autocorrelation, but some areas in toronto tends to have higher lead concentrations than the rest."
 format: pdf
@@
 risks, particularly for vulnerable populations such as children and
-pregnant women(Government Canada).
+pregnant women(@c-1).
 
@@
 
-While significant research has extensively documented the public health
+While significant research has documented the public health
 impacts of lead exposure, particularly focusing on its toxic effects on
-vulnerable populations like children and pregnant women (Levallois et
-al., 2018)(Jarvis & Fawell, 2021), a critical gap remains in
+vulnerable populations like children and pregnant women (@levallois2018public)(@jarvis2021lead), a critical gap remains in
 understanding how lead concentrations in urban drinking water systems
@@
 on previous lead levels, and there is insufficient evidence to suggest
-that certain months consistently exhibit higher trends of lead
-concentrations compared to others.
+that certain months or years consistently have higher trends of lead
+concentrations.
 
-However, some district does have higher trend of lead concentration.
-There are clusters of outliers within some specific district and some
-district have higher mean lead concentrations than other. It is safe to
+However, some districts do have higher trend of lead concentration.
+There are clusters of outliers within some specific districts and some
+districts have higher mean lead concentrations than others. It is safe to
 say that geographic locations and community water maintenance have an
@@
 
-This topic is important because even low levels of lead exposure can
+This topic is significant because even low levels of lead exposure can
 have serious health consequences, including developmental delays,
-cognitive impairments and even detrimental effects(Government Canada).
+cognitive impairments and even detrimental effects(@c-1).
 
-We use @talia for data analysis, @opendatatorotno for collected data
-set, and @rohan for directions and template.
+@talia is used for data analysis, @opendatatorotno for collected data
+set, and @alexander2023telling for instructions and template.
 
 The remainder of this paper is structured as follows. @sec-mydatasection
-Data, further explains the data set we are working with and the
-prepossessing done for further analysis. A discussion with the
-measurement used when collecting the data is also included. Graphs and
-tables were used to show important features of the data set.
-@sec-mydiscussion Discussion, talks about, firstly, to what extent will
+Data, provides a detailed explanation of the data set working with and the prepossessing done for further analysis. A discussion with the
+measurement used when collecting the data is also included in @sec-measurement. Graphs and tables were used to emphasis the importance of the features in the data set.
+@sec-mydiscussion Discussion, covers, @sec-health, to what extent will
 Lead Concentration influence Health and how much concentration will be
-categorized as “dangerous”. Secondly, the paper went into a detailed
+categorized as “dangerous”. In @sec-district, the paper went into a detailed
 analysis of how districts affect lead concentration. Graphs and outliers
-analysis was used to present the result. Thirdly, the paper performs
+analysis was used to present the result. Thirdly in @sec-time, the paper performs
 Times Series analysis to analyze the lead concentration over time.
-Lastly, there is a subsection discussing weakness of the paper and some
+Lastly, @sec-future discussed weakness of the paper and some
 further steps to further support the question.
@@
 
-#TODO: change library citation R (R Core Team 2023) was used for the
-data analysis and the productions of all graphs and tables. The
-libraries tidyverse (Wickham et al. 2019), here (J. B. Kirill Müller
-2020), dplyr (Hadley Wickham 2023), tibble (R. F. Kirill Müller Hadley
-Wickham 2023), janitor (Sam Firke 2023), ggplot2 (Wickham 2016), and
-knitr (Xie 2023) were also used for organizing data and performing
-analysis.
+R (R Core Team 2023) was used for the data analysis and the productions of all graphs and tables.
+The libraries: tidyverse (@o), dplyr (Hadley Wickham 2023), ggplot2 (@ggplot2), gt(@gt) , lubridtate(@lubridate), tidyr(@tidyr), and knitr (@knitr) were also used for organizing data and performing analysis.
 
 ## Data Set and Context
-The data set "Non Regulated Lead Sample" by Open Data Toronto is to
+The data set "Non Regulated Lead Sample" by Open Data Toronto aims to
 address the concerns Toronto residents have for their drinking water.
 The data set contains information on lead concentrations in water
-samples. Each row is from a water sample, with corresponding information
+samples. Each row represents a water sample, with corresponding information
 on the date, partial postal code where the sample was provided, and the
@@
 
-Lead exposure is an important public health issue. High levels of lead
+Lead exposure is an significant public health issue. High levels of lead
 in drinking water can pose severe health risks, particularly to
-vulnerable populations.(WHO) Chronic exposure to lead can result in
-developmental delays, neurological damage, and a range of other health
-problems.(WHO) Tracking lead concentrations in water is crucial for
+vulnerable populations.(@w) Chronic exposure to lead can result in
+developmental delays, neurological damage, and varous of other health
+problems.(@w) Monitoring lead concentrations in water is crucial for
 ensuring safe water quality and identifying areas requiring intervention
@@
 
-There are similar similar data sets that addresses the problem of hard
-water contamination. One example is in state of Michigan, USA, MIDHHS
+There are similar data sets that address the problem of hard
+water contamination. One example is from state of Michigan, USA, MIDHHS
 collected similar sample with months, years and lead
-concentrations(MIDHHS). There are also other data set that analyze a
-different metal in water. However, this data set is used because lead is
-a very common metal in hard water with severe damage to public health
-when overdose, and being a resident in city of Toronto, this should be a
-concern that needs attention.
+concentrations(@m). Other data setanalyze the
+different metals in water. However, this data set was chosen because lead is
+a common metal in hard water with severe damage to public health
+when overdose, and being a resident in city of Toronto, this is a
+concern that needs to raise attention .
 
@@
 
-There are five variables in total: \_id, Sample Number, Sample Date,
+The five variables in this paper: \_id, Sample Number, Sample Date,
 PartialPostalCode and Lead Amount. While the first two are keys of this
-data set for identification purpose, the rest are our variable of
-interest:
+data set for identification purpose, the remaining varaiables are of our interest to our analysis:
 
@@
 can be used to detect trends and seasonal variations.
+It is evident from (@fig-showdate) that the data points are evenly distributed through time except during 2020, there is an empty period of time.
+This might be caused by COVID-19 quarantine, where the residence does not have the chance to volunteer their water sample.
+```{r}
+#| label: fig-showdate
+#| fig-cap: "Graph of Dates Categorized by Year and Presented by Each Months"
+#| echo: false
+# Create a new data frame with extracted year and month
+data_date <- data %>%
+  transmute(
+    Year = year(`Sample Date`),
+    Month = month(`Sample Date`, label = TRUE, abbr = TRUE)
+  )
+
+# Plot with ggplot2
+ggplot(data_date, aes(x = Year, y = Month)) +
+  geom_jitter(width = 0.2, height = 0.2, color = blueishgrey, size = 3) + # Add jitter to separate points
+  labs(title = "Distribution of Sample Dates by Year and Month",
+       x = "Year",
+       y = "Month") +
+  theme_minimal()
+
+```
 
@@
 identify potential geographical clusters.
+The (@fig-splitpostal) shows the partial postal code categorized by the first two characters and the y axis being the last character ranging from A-Z.
+It can be oberseved from (@fig-splitpostal) that the data points are not evenly distributed.
+This can be caused by sampling bias since it is a volunterring sampling but it can also be caused by the fact that some areas in City of Toronto are not present for residency.
+```{r}
+#| label: fig-splitpostal
+#| fig-cap: "Graph of Partial Postal Code Splitting into First Two Characters and Last Character"
+#| echo: false
+
+# Create a new data frame with extracted components
+data_postal <- data %>%
+  transmute(
+    FirstTwoChars = substr(district, 1, 2),
+    LastChar = substr(district, 3, 3)
+  )
+
+# Plot the extracted components using ggplot2
+ggplot(data_postal, aes(x = FirstTwoChars, y = LastChar)) +
+  geom_jitter(width = 0.2, height = 0.2, color = blueishgrey, size = 3) + # Jitter points to spread them out for better visualization
+  labs(title = "Relationship Between First Two and Last Characters of Postal Codes",
+       x = "First Two Characters",
+       y = "Last Character") +
+  theme_minimal()
+
+```
 
 **Lead Concentration (ppm):** The measured amount of lead in the water
-sample, expressed in parts per million. This is the primary variable of
+sample, expressed in parts per million(@fig-boxplot). This is the primary variable of
 interest, as it directly relates to public health risks associated with
-lead exposure. We can also further analyze combinations of lead
-concentration with other variables.
+lead exposure.Additionally,     combinations of lead concentration with other variables can be furthur analyzed.
+```{r}
+#| label: fig-boxplot
+#| fig-cap: "Boxplot showing the lead concentration above the saftey threshold"
+#| echo: false
+
+# Set the threshold value for the horizontal line
+threshold_value <- 0.010011423
+
+# Create the boxplot using ggplot2
+ggplot(data, aes(y = `Lead Amount (ppm)`)) +
+  geom_boxplot(fill = "skyblue") +
+  geom_hline(yintercept = threshold_value, color = "red", linetype = "dashed") +
+  labs(title = "Boxplot of Lead Amount (ppm) with Threshold Line",
+       y = "Lead Amount (ppm)") +
+  theme_minimal()
+
+```
 
@@
 ## Summary
-Table 1 is a summary table for num type variable lead amount.
-We can see that the minimum and maximum value differs by 168,800 times, which is very significant.
-Therefore, we can assume that there are water samples that are very clean in terms of lead concentration and there are samples that are contaminated.
-We can also see that the mean and median differs a lot.
+A summary table(@tbl-summary) for num type variable lead amount.It is evident that the minimum and maximum value differs by 168,800 times, which is very significant.
+In (@gif-boxplot), it is evident that there are many outliers.
+Therefore, the assumption can be made that there are water samples that are very clean in terms of lead concentration and there are samples that are contaminated.
+It is clear that the mean and median differs a lot.
 Which means that the distribution of this data set is highly right skewed.
-It also shows evidence that there are many outliers that will elevated mean lead concentration level in this data set.
+The outliers will elevated mean lead concentration level in this data set.
 
@@
 #| label: tbl-summary
-#| tbl-cap: "Distribution Statistics Measurements of Lead Amount"
+#| tbl-cap: "Distribution Statistics Measurements of Lead Amount in ppm"
 #| echo: false
@@
 
-## Measurement
-In this dataset, we are interested in the concentration of lead in drinking water across different regions of a city and over time.
+## Measurement{#sec-measurement}
+In this dataset, the aim is to explore in the concentration of lead in drinking water across different regions of a city and over time.
 The measurement process transforms a real-world phenomenon—lead contamination in water—into quantitative entries in a dataset that can be analyzed for trends.
@@
 
-## Lead Concentration influencing Health
+## Lead Concentration influencing Health{#sec-health}
 
@@
 
-The main variable we are examining is the Lead Amount in
+The main variable examining in this paper is the Lead Amount in
 ppm. The distribution of the data is highly right skewed with most of
@@
 and 4 of them ranging from 3 to 8.44 which is the maximum value as shown
-in the summary Table 1. We can also verify our observation
+in the summary Table 1. Additionally observation can be verified
 since the summary Table 1 shows that 75% of the data is below
 0.001480. Government of Canada had specified that the maximum acceptable concentration (MAC) of drinking
-water is 0.010mg/L(Government Canada), which converts to 0.010011423
-part/million(unitconverters.net), this is shown by the red line in the
+water is 0.010mg/L(@c-1), which converts to 0.010011423
+part/million(@c), this is shown by the red line in the
 figure 1(@fig-boxplot).
-```{r}
-#| label: fig-boxplot
-#| fig-cap: Boxplot showing the lead concentration above the saftey threshold
-#| echo: false
-
-# Set the threshold value for the horizontal line
-threshold_value <- 0.010011423
-
-# Create the boxplot using ggplot2
-ggplot(data, aes(y = `Lead Amount (ppm)`)) +
-  geom_boxplot(fill = "skyblue") +
-  geom_hline(yintercept = threshold_value, color = "red", linetype = "dashed") +
-  labs(title = "Boxplot of Lead Amount (ppm) with Threshold Line",
-       y = "Lead Amount (ppm)") +
-  theme_minimal()
-
-```
 
@@
 As shown in (@tbl-threshold), there are only 280 counts of water sample collected that is above MAC, which takes up only 2.74% of the whole data set.
-Even though this seems like a good result, those 280 measurements still effects people and their family's health since those lead concentrations will have negative effect on health(WHO)(Canadian Government).
+Even though this seems like a good result, those 280 measurements still effects people and their family's health since those lead concentrations will have negative effect on health(@w)(@c-1).
 ```{r, echo=FALSE}
 #| label: tbl-threshold
-#| tbl-cap: "Number and Percentage of data above MAC"
+#| tbl-cap: "Number and Percentage of data above Maximum Acceptable Concentration"
 #| echo: false
@@
 ### plot of the data above this threshold
-To visualize those data above MAC, data below MAC and above 2 were filtered out, because from (@fig-boxplot), we can see that there are only 4 data points above 2, therefore, with that in mind it is safe to temporarily ignore to get a scale that will better show the most data points above MAC.
+To visualize those data above MAC, data below MAC and above 2 were filtered out, because from (@fig-boxplot), it is clear that there are only 4 data points above 2, therefore, with that in mind it is safe to temporarily ignore to get a scale that will better show the most data points above MAC.
 A scatterplot was used with x axis being date and y axis being lead concentration give a graph that spreads out lead concentrations instead of on a 1-dimensional line.
-From (@fig-between) we can see that there are clusters of data points below 0.25 and for higher lead concentrations, there are less data points and more spread out than those below 0.25.
+From (@fig-between) this indicates that there are clusters of data points below 0.25 and for higher lead concentrations, there are less data points and more spread out than those below 0.25.
 
@@
 #| label: fig-between
-#| fig-cap: "Data between 2 and MAC representing with Date"
+#| fig-cap: "Data Between 2ppm and Maximum Acceptable Concentration Representing By Date"
 #| echo: false
@@
 
-## How districts affect Lead Concentration
+## How districts affect Lead Concentration{#sec-district}
 
@@
 The districts containing outliers were also shown in (@fig-boxplotdistrict).
-We can see that the "M4" has two data points above 8 ppm, as
+It is clear that the "M4" has two data points above 8 ppm, as
 well as a cluster between 0 and 1 ppm. "M6" also have many data between 0 and
@@
 #| label: fig-boxplotdistrict
-#| fig-cap: "Boxplot Showing Outliers categorized by district"
+#| fig-cap: "Boxplot Showing Outliers Categorized by District"
 #| echo: false
@@
 
-We will verify which district have a higher overall lead concentration by calculating the mean lead concentration when still categorize the partial postal codes by only the
+Verification will be conducted which district have a higher overall lead concentration by calculating the mean lead concentration when still categorize the partial postal codes by only the
 first two numbers. Since outliers might influence the mean, especially
@@
 concentration without outliers above 2 ppm and categorized by only the
-first two characters of the partial postal code. We can see that the top
+first two characters of the partial postal code. It is evident that the top
 two district of the highest lead concentration in water is M9 and M6 respectively.
@@
 #| label: tbl-meandistrict
-#| tbl-cap: "Specific Numbers of the mean for each district"
+#| tbl-cap: "Specific Numbers of the Mean Lead Concentration for each District"
 #| echo: false
@@
 ```
-
+\newpage
 ```{r}
 #| label: fig-meandistrict
-#| fig-cap: "Mean lead concentration categorized by first two characters in partial postal code"
+#| fig-cap: "Mean Lead Concentration Categorized by First Two Characters from Partial Postal Code"
 #| echo: false
@@
 
-## How months and year affect Lead Concentration
+## How months and year affect Lead Concentration{#sec-time}
 
@@
 
-In this dataset, we have variables time and lead amount,
-therefore, we can perform times series analysis. As shown in
-(@fig-timeseries), we can see that except for some sharp increase
+In this dataset, the variables are time and lead amount,
+therefore,times series analysis can be performed. As shown in
+(@fig-timeseries), it is evident that except for some sharp increase
 in 2014, 2019, 2021, 2023, the overall trend is stays around 0 and
@@
 
-From the ACF and PACF figures, we can conclude that the previous values
+From the ACF and PACF figures, it can be concluded that the previous values
 of lead concentrations does not impact future values. For ACF, when lag
@@
 
-Lead Concentration in water can vary overtime(WHO), therefore the high lead concentration outliers might not be caused by correlation instead, "a probability-based adaptive sampling plan should be used to access exposure(WHO)".
+Lead Concentration in water can vary overtime(@w), therefore the high lead concentration outliers might not be caused by correlation instead, "a probability-based adaptive sampling plan should be used to access exposure(@w)".
 
@@
 #| label: fig-timeseries
-#| fig-cap: "Lead Levels changed over time for times seriese analysis"
+#| fig-cap: "Lead Levels Changed Over Time for Times Series Analysis"
 #| echo: false
@@
 #| label: fig-ACF
-#| fig-cap: "Autocorrelation function for lead concentrations"
+#| fig-cap: "Autocorrelation Function for Lead Concentrations"
 #| echo: false
@@
 #| label: fig-PACF
-#| fig-cap: "Partial Autocorrelation function for lead concentrations"
+#| fig-cap: "Partial Autocorrelation Function for Lead Concentrations"
 #| echo: false
@@
 ### graphs of months vs mean lead concentration
-With all data points, we do not see obvious trends over time.
-Here we examine closely for the constructed variable month from Date.
+With all data points, there are no obvious trends over time.
+The examination is made closely for the constructed variable month from Date.
 (@fig-month) is the boxplot showing the outliers, but there are not enough evidence to show that there are specific month that will have a higher lead concentration.
-Therefore, we perform a similar filtering on data that is larger than 2 ppm.
+Therefore, a similar filtering is applied to data larger than 2 ppm.
 ```{r}
 #| label: fig-month
-#| fig-cap: "Boxplots that shows the outliers categorized by Months"
+#| fig-cap: "Boxplots that Shows the Outliers Categorized by Months"
 #| echo: false
@@
 However, this graph does not provide enough evidence to prove seasonality.
-We need we examine whether April and September for every year will repeatedly have higher measured lead concentration.
+It is necessary to examine whether April and September for every year will repeatedly have higher measured lead concentration.
 This will be further discussed in @sec-future.
@@
 #| label: fig-meanmonth
-#| fig-cap: "Mean Lead Concentration Without Data Greater than 2 ppm categorized by Months"
+#| fig-cap: "Mean Lead Concentration Without Data Greater than 2 ppm Categorized by Months"
 #| echo: false
@@
 ### graphs of years vs mean lead concentration
-Similar to months, we also constructed a variable call year.
+Similar to months,variable called year was also constructed.
 However, both the boxplots and the mean values without data below 2 ppm does not show trend from year to year.
@@
 #| label: fig-year
-#| fig-cap: "Boxplots that shows the outliers categorized by Year"
+#| fig-cap: "Boxplots that Shows the Outliers Categorized by Year"
 #| echo: false

### 2024-09-27 scripts/00-simulate_data.R
fixed error in simulated data

@@ -39,14 +39,8 @@ data <-
     sample_number = sample_number,
     sample_date = sample_date,
     partial_postal_code = partial_postal_code,
-    dates = as.Date(
-      runif(
-        n = number_of_dates,
-        min = as.numeric(start_date),
-        max = as.numeric(end_date)
-      ),
-      origin = "2014-01-01"
-    ),
+
+    
     # used gamma distibution instead of the poison distribution because the mean_lead is a continuous variable,
     # if we use poisson distibution, all simulations will be 0
     leadAmountppm = rgamma(n = number_of_dates, shape = 2, rate = 1 / mean_lead)

### 2024-09-27 paper/paper.qmd
final formatting of the paper

@@
 
-While significant research has documented the public health
-impacts of lead exposure, particularly focusing on its toxic effects on
-vulnerable populations like children and pregnant women (@levallois2018public)(@jarvis2021lead), a critical gap remains in
+While significant research has documented the public health impacts of
+lead exposure, particularly focusing on its toxic effects on vulnerable
+populations like children and pregnant women
+(@levallois2018public)(@jarvis2021lead), a critical gap remains in
 understanding how lead concentrations in urban drinking water systems
@@
 There are clusters of outliers within some specific districts and some
-districts have higher mean lead concentrations than others. It is safe to
-say that geographic locations and community water maintenance have an
+districts have higher mean lead concentrations than others. It is safe
+to say that geographic locations and community water maintenance have an
 effect on water safety.
@@
 The remainder of this paper is structured as follows. @sec-mydatasection
-Data, provides a detailed explanation of the data set working with and the prepossessing done for further analysis. A discussion with the
-measurement used when collecting the data is also included in @sec-measurement. Graphs and tables were used to emphasis the importance of the features in the data set.
-@sec-mydiscussion Discussion, covers, @sec-health, to what extent will
-Lead Concentration influence Health and how much concentration will be
-categorized as “dangerous”. In @sec-district, the paper went into a detailed
-analysis of how districts affect lead concentration. Graphs and outliers
-analysis was used to present the result. Thirdly in @sec-time, the paper performs
-Times Series analysis to analyze the lead concentration over time.
-Lastly, @sec-future discussed weakness of the paper and some
-further steps to further support the question.
+Data, provides a detailed explanation of the data set working with and
+the prepossessing done for further analysis. A discussion with the
+measurement used when collecting the data is also included in
+@sec-measurement. Graphs and tables were used to emphasis the importance
+of the features in the data set. @sec-mydiscussion Discussion, covers,
+@sec-health, to what extent will Lead Concentration influence Health and
+how much concentration will be categorized as “dangerous”. In
+@sec-district, the paper went into a detailed analysis of how districts
+affect lead concentration. Graphs and outliers analysis was used to
+present the result. Thirdly in @sec-time, the paper performs Times
+Series analysis to analyze the lead concentration over time. Lastly,
+@sec-future discussed weakness of the paper and some further steps to
+further support the question.
 
@@
 
-R (R Core Team 2023) was used for the data analysis and the productions of all graphs and tables.
-The libraries: tidyverse (@o), dplyr (Hadley Wickham 2023), ggplot2 (@ggplot2), gt(@gt) , lubridtate(@lubridate), tidyr(@tidyr), and knitr (@knitr) were also used for organizing data and performing analysis.
+R (R Core Team 2023) was used for the data analysis and the productions
+of all graphs and tables. The libraries: tidyverse (@o), dplyr (Hadley
+Wickham 2023), ggplot2 (@ggplot2), gt(@gt) , lubridtate(@lubridate),
+tidyr(@tidyr), and knitr (@knitr) were also used for organizing data and
+performing analysis.
 
 ## Data Set and Context
+
 The data set "Non Regulated Lead Sample" by Open Data Toronto aims to
@@
 The data set contains information on lead concentrations in water
-samples. Each row represents a water sample, with corresponding information
-on the date, partial postal code where the sample was provided, and the
-main variable of interest, lead concentration in parts per million
-(ppm).
+samples. Each row represents a water sample, with corresponding
+information on the date, partial postal code where the sample was
+provided, and the main variable of interest, lead concentration in parts
+per million (ppm).
 
@@
 
-There are similar data sets that address the problem of hard
-water contamination. One example is from state of Michigan, USA, MIDHHS
-collected similar sample with months, years and lead
-concentrations(@m). Other data setanalyze the
-different metals in water. However, this data set was chosen because lead is
-a common metal in hard water with severe damage to public health
-when overdose, and being a resident in city of Toronto, this is a
-concern that needs to raise attention .
+There are similar data sets that address the problem of hard water
+contamination. One example is from state of Michigan, USA, MIDHHS
+collected similar sample with months, years and lead concentrations(@m).
+Other data setanalyze the different metals in water. However, this data
+set was chosen because lead is a common metal in hard water with severe
+damage to public health when overdose, and being a resident in city of
+Toronto, this is a concern that needs to raise attention .
 
@@
 PartialPostalCode and Lead Amount. While the first two are keys of this
-data set for identification purpose, the remaining varaiables are of our interest to our analysis:
+data set for identification purpose, the remaining varaiables are of our
+interest to our analysis:
 
 **Date**: Represents the date on which the water sample was collected,
-can be used to detect trends and seasonal variations.
-It is evident from (@fig-showdate) that the data points are evenly distributed through time except during 2020, there is an empty period of time.
-This might be caused by COVID-19 quarantine, where the residence does not have the chance to volunteer their water sample.
+can be used to detect trends and seasonal variations. It is evident from
+(@fig-showdate) that the data points are evenly distributed through time
+except during 2020, there is an empty period of time. This might be
+caused by COVID-19 quarantine, where the residence does not have the
+chance to volunteer their water sample.
+
 ```{r}
@@
 the general area where the water sample was taken, can be used to
-identify potential geographical clusters.
-The (@fig-splitpostal) shows the partial postal code categorized by the first two characters and the y axis being the last character ranging from A-Z.
-It can be oberseved from (@fig-splitpostal) that the data points are not evenly distributed.
-This can be caused by sampling bias since it is a volunterring sampling but it can also be caused by the fact that some areas in City of Toronto are not present for residency.
+identify potential geographical clusters. The (@fig-splitpostal) shows
+the partial postal code categorized by the first two characters and the
+y axis being the last character ranging from A-Z. It can be oberseved
+from (@fig-splitpostal) that the data points are not evenly distributed.
+This can be caused by sampling bias since it is a volunterring sampling
+but it can also be caused by the fact that some areas in City of Toronto
+are not present for residency.
+
 ```{r}
@@
   geom_jitter(width = 0.2, height = 0.2, color = blueishgrey, size = 3) + # Jitter points to spread them out for better visualization
-  labs(title = "Relationship Between First Two and Last Characters of Postal Codes",
+  labs(title = "Relationship Between Splitted Postal Codes",
        x = "First Two Characters",
@@
 **Lead Concentration (ppm):** The measured amount of lead in the water
-sample, expressed in parts per million(@fig-boxplot). This is the primary variable of
-interest, as it directly relates to public health risks associated with
-lead exposure.Additionally,     combinations of lead concentration with other variables can be furthur analyzed.
+sample, expressed in parts per million(@fig-boxplot). This is the
+primary variable of interest, as it directly relates to public health
+risks associated with lead exposure.Additionally, combinations of lead
+concentration with other variables can be furthur analyzed.
+
 ```{r}
@@
 ggplot(data, aes(y = `Lead Amount (ppm)`)) +
-  geom_boxplot(fill = "skyblue") +
+  geom_boxplot(fill = "skyblue",outlier.colour = "red") +
   geom_hline(yintercept = threshold_value, color = "red", linetype = "dashed") +
@@
 For the purpose of data analysis the raw data set went through some
-pre-processing: NA Values: All missing (NA) values were removed to
-ensure accurate analysis. Lead Concentration: The Lead Amount (ppm)
-variable was cleaned to remove any anomalies and ensure that values were
-properly numeric. Date Variable Type: The Date variable was converted to
-Date type in R to facilitate temporal analysis and aggregation by months
-and years.
+pre-processing:
+
+**NA Values:** All missing (NA) values were removed to ensure accurate
+analysis.
+
+**Lead Concentration:** The Lead Amount (ppm) variable was cleaned to
+remove any anomalies and ensure that values were properly numeric.
+
+**Date Variable Type:** The Date variable was converted to Date type in
+R to facilitate temporal analysis and aggregation by months and years.
 
@@
 ## Summary
-A summary table(@tbl-summary) for num type variable lead amount.It is evident that the minimum and maximum value differs by 168,800 times, which is very significant.
-In (@gif-boxplot), it is evident that there are many outliers.
-Therefore, the assumption can be made that there are water samples that are very clean in terms of lead concentration and there are samples that are contaminated.
-It is clear that the mean and median differs a lot.
-Which means that the distribution of this data set is highly right skewed.
-The outliers will elevated mean lead concentration level in this data set.
+
+A summary table(@tbl-summary) for num type variable lead amount.It is
+evident that the minimum and maximum value differs by 168,800 times,
+which is very significant. In (@fig-boxplot), it is evident that there
+are many outliers. Therefore, the assumption can be made that there are
+water samples that are very clean in terms of lead concentration and
+there are samples that are contaminated. It is clear that the mean and
+median differs a lot. Which means that the distribution of this data set
+is highly right skewed. The outliers will elevated mean lead
+concentration level in this data set.
 
@@
 ```
+
 ## Relationships
+
 These relationships are upon further discussion in @sec-mydiscussion.
@@
 
-## Measurement{#sec-measurement}
-In this dataset, the aim is to explore in the concentration of lead in drinking water across different regions of a city and over time.
-The measurement process transforms a real-world phenomenon—lead contamination in water—into quantitative entries in a dataset that can be analyzed for trends.
+## Measurement {#sec-measurement}
 
-**Lead concentration:** Measured in parts per million (ppm), indicating the amount of lead in the water relative to its volume.
-This choice of unit is small enough to capture low but potentially harmful concentrations and is standard in water quality testing.
-This transformation from the chemical presence of lead to a numeric value is a key aspect of measurement.
+In this dataset, the aim is to explore in the concentration of lead in
+drinking water across different regions of a city and over time. The
+measurement process transforms a real-world phenomenon—lead
+contamination in water—into quantitative entries in a dataset that can
+be analyzed for trends.
 
-**Sample Date:** The date the sample was taken is recorded to allow for analysis over time, detecting patterns.
+**Lead concentration:** Measured in parts per million (ppm), indicating
+the amount of lead in the water relative to its volume. This choice of
+unit is small enough to capture low but potentially harmful
+concentrations and is standard in water quality testing. This
+transformation from the chemical presence of lead to a numeric value is
+a key aspect of measurement.
 
-**Partial Postal Code:** Postal code represents where the sample was collected, allowing comparisons between different geographic areas.
-This made it easy to get a result using data analysis to public health control organizations.
+**Sample Date:** The date the sample was taken is recorded to allow for
+analysis over time, detecting patterns.
 
-Some challenges are that this is a non-regulated program, therefore the measurement process is not controlled, decreasing accuracy and precision.
-However, this is minimized by providing a detailed instructions with the sample kit used to collect the water samples and the deviation will only affect Lead concentration.
-Date and district are accurate.
+**Partial Postal Code:** Postal code represents where the sample was
+collected, allowing comparisons between different geographic areas. This
+made it easy to get a result using data analysis to public health
+control organizations.
 
-Choices about measurement such as the instruments used, units of measurement, and protocols for sample collection—affect how well the data set represents the real-world phenomenon of lead contamination.
+Some challenges are that this is a non-regulated program, therefore the
+measurement process is not controlled, decreasing accuracy and
+precision. However, this is minimized by providing a detailed
+instructions with the sample kit used to collect the water samples and
+the deviation will only affect Lead concentration. Date and district are
+accurate.
+
+Choices about measurement such as the instruments used, units of
+measurement, and protocols for sample collection—affect how well the
+data set represents the real-world phenomenon of lead contamination.
 
@@
 
-## Lead Concentration influencing Health{#sec-health}
+## Lead Concentration Influencing Health {#sec-health}
 
-### WHO and Canadian federal guidance on lead concentration in water
+### WHO and Canadian Federal Guidance on Lead Concentration in Water
 
-The main variable examining in this paper is the Lead Amount in
-ppm. The distribution of the data is highly right skewed with most of
-its data being very small, as shown in (@fig-boxplot), most
-data are almost 0 with outliers which are mostly less than 2 ppm
-and 4 of them ranging from 3 to 8.44 which is the maximum value as shown
-in the summary Table 1. Additionally observation can be verified
-since the summary Table 1 shows that 75% of the data is below
-0.001480. Government of Canada had specified that the maximum acceptable concentration (MAC) of drinking
-water is 0.010mg/L(@c-1), which converts to 0.010011423
-part/million(@c), this is shown by the red line in the
-figure 1(@fig-boxplot).
+The main variable examining in this paper is the Lead Amount in ppm. The
+distribution of the data is highly right skewed with most of its data
+being very small, as shown in (@fig-boxplot), most data are almost 0
+with outliers which are mostly less than 2 ppm and 4 of them ranging
+from 3 to 8.44 which is the maximum value as shown in the summary
+Table 1. Additionally observation can be verified since the summary
+Table 1 shows that 75% of the data is below 0.001480. Government of
+Canada had specified that the maximum acceptable concentration (MAC) of
+drinking water is 0.010mg/L(@c-1), which converts to 0.010011423
+part/million(@c), this is shown by the red line in the figure
+1(@fig-boxplot).
 
-### Percentage of data above this threshold
+### Percentage of Data Above the Threshold
 
-From calculations by R, there are 97.26% of the data collected indicating that
-the lead amount in the water is safe.
-As shown in (@tbl-threshold), there are only 280 counts of water sample collected that is above MAC, which takes up only 2.74% of the whole data set.
-Even though this seems like a good result, those 280 measurements still effects people and their family's health since those lead concentrations will have negative effect on health(@w)(@c-1).
+From calculations by R, there are 97.26% of the data collected
+indicating that the lead amount in the water is safe. As shown in
+(@tbl-threshold), there are only 280 counts of water sample collected
+that is above MAC, which takes up only 2.74% of the whole data set. Even
+though this seems like a good result, those 280 measurements still
+effects people and their family's health since those lead concentrations
+will have negative effect on health(@w)(@c-1).
+
 ```{r, echo=FALSE}
@@
 
-### plot of the data above this threshold
-To visualize those data above MAC, data below MAC and above 2 were filtered out, because from (@fig-boxplot), it is clear that there are only 4 data points above 2, therefore, with that in mind it is safe to temporarily ignore to get a scale that will better show the most data points above MAC.
-A scatterplot was used with x axis being date and y axis being lead concentration give a graph that spreads out lead concentrations instead of on a 1-dimensional line.
-From (@fig-between) this indicates that there are clusters of data points below 0.25 and for higher lead concentrations, there are less data points and more spread out than those below 0.25.
+### Plot of the Data Above the Threshold
+
+To visualize those data above MAC, data below MAC and above 2 were
+filtered out, because from (@fig-boxplot), it is clear that there are
+only 4 data points above 2, therefore, with that in mind it is safe to
+temporarily ignore to get a scale that will better show the most data
+points above MAC. A scatterplot was used with x axis being date and y
+axis being lead concentration give a graph that spreads out lead
+concentrations instead of on a 1-dimensional line. From (@fig-between)
+this indicates that there are clusters of data points below 0.25 and for
+higher lead concentrations, there are less data points and more spread
+out than those below 0.25.
 
@@
 
-## How districts affect Lead Concentration{#sec-district}
+## How Districts Sffect Lead Concentration {#sec-district}
 
-### graphs of postal code vs lead concentrations
+### Graphs of Postal Code vs Lead Concentrations
 
-Nine boxplot (@fig-boxplotdistrict) each representing a district, which is the constructed variable from the first two letters of partial postal code.
-Again, most data points are around 0, and the outliers streched the scale so that this way we can easily analyze the outliers.
-The districts containing outliers were also shown in (@fig-boxplotdistrict).
-It is clear that the "M4" has two data points above 8 ppm, as
-well as a cluster between 0 and 1 ppm. "M6" also have many data between 0 and
-2ppm. It is safe to assume that these two districts have water sources with high lead concentrations.
+Nine boxplot (@fig-boxplotdistrict) each representing a district, which
+is the constructed variable from the first two letters of partial postal
+code. Again, most data points are around 0, and the outliers streched
+the scale so that this way we can easily analyze the outliers. The
+districts containing outliers were also shown in (@fig-boxplotdistrict).
+It is clear that the "M4" has two data points above 8 ppm, as well as a
+cluster between 0 and 1 ppm. "M6" also have many data between 0 and
+2ppm. It is safe to assume that these two districts have water sources
+with high lead concentrations.
+
 ```{r}
@@
 
-### extract outliers above the threshold
+### Extract Outliers Above the Threshold
 
-Verification will be conducted which district have a higher overall lead concentration by calculating the mean lead concentration when still categorize the partial postal codes by only the
-first two numbers. Since outliers might influence the mean, especially
-when in our dataset we are dealing with median 0.000473 and outliers as
-high as 8ppm, the outliers above 2 ppm were omitted before caculating the mean lead concentration.
+Verification will be conducted which district have a higher overall lead
+concentration by calculating the mean lead concentration when still
+categorize the partial postal codes by only the first two numbers. Since
+outliers might influence the mean, especially when in our dataset we are
+dealing with median 0.000473 and outliers as high as 8ppm, the outliers
+above 2 ppm were omitted before caculating the mean lead concentration.
 
-### mean value of lead concentration for different districts
+### Mean Values of Lead Concentration for Different Districts
 
@@
 concentration without outliers above 2 ppm and categorized by only the
-first two characters of the partial postal code. It is evident that the top
-two district of the highest lead concentration in water is M9 and M6 respectively.
-This shows that without the data above 2 ppm, M4 does not have a overall high lead concentration, but only some very high outliers.
-However, M9 and M6 will be the two most dangerous district in Toronto in terms of lead concentration in water.
-Even though all the mean values shown in (@tbl-meandistrict) are all below MAC by Canadian Government, the M9 and M6 district might need more attention in the topic of water saftey given that the rest districts perform better especially M8.
+first two characters of the partial postal code. It is evident that the
+top two district of the highest lead concentration in water is M9 and M6
+respectively. This shows that without the data above 2 ppm, M4 does not
+have a overall high lead concentration, but only some very high
+outliers. However, M9 and M6 will be the two most dangerous district in
+Toronto in terms of lead concentration in water. Even though all the
+mean values shown in (@tbl-meandistrict) are all below MAC by Canadian
+Government, the M9 and M6 district might need more attention in the
+topic of water saftey given that the rest districts perform better
+especially M8.
 
@@
 ```
+
 \newpage
+
 ```{r}
@@
 
-## How months and year affect Lead Concentration{#sec-time}
+## How Months and Years Affect Lead Concentration {#sec-time}
 
-### Times Series analysis for autocorrelation{#sec-timeseries}
+### Times Series Analysis for Autocorrelation {#sec-timeseries}
 
-In this dataset, the variables are time and lead amount,
-therefore,times series analysis can be performed. As shown in
-(@fig-timeseries), it is evident that except for some sharp increase
-in 2014, 2019, 2021, 2023, the overall trend is stays around 0 and
-remains positive. There are no increasing or decreasing trend overtime and no obvious seasonality present.
+In this dataset, the variables are time and lead amount, therefore,times
+series analysis can be performed. As shown in (@fig-timeseries), it is
+evident that except for some sharp increase in 2014, 2019, 2021, 2023,
+the overall trend is stays around 0 and remains positive. There are no
+increasing or decreasing trend overtime and no obvious seasonality
+present.
 
-From the ACF and PACF figures, it can be concluded that the previous values
-of lead concentrations does not impact future values. For ACF, when lag
-is 0, ACF is always 1 but for all other lags in ACF(@fig-ACF) and
-PACF(@fig-PACF) they are all within the dotted blue lines which is
-the evidence that there are no significant correlation at any lags present in the data.
+From the ACF and PACF figures, it can be concluded that the previous
+values of lead concentrations does not impact future values. For ACF,
+when lag is 0, ACF is always 1 but for all other lags in ACF(@fig-ACF)
+and PACF(@fig-PACF) they are all within the dotted blue lines which is
+the evidence that there are no significant correlation at any lags
+present in the data.
 
@@
 incident of a very high lead concentration might again only be outliers
-instead of due some factors that will effect the overall water quality in
-Toronto.
+instead of due some factors that will effect the overall water quality
+in Toronto.
 
-Lead Concentration in water can vary overtime(@w), therefore the high lead concentration outliers might not be caused by correlation instead, "a probability-based adaptive sampling plan should be used to access exposure(@w)".
+Lead Concentration in water can vary overtime(@w), therefore the high
+lead concentration outliers might not be caused by correlation instead,
+"a probability-based adaptive sampling plan should be used to access
+exposure(@w)".
 
@@
 
-### graphs of months vs mean lead concentration
-With all data points, there are no obvious trends over time.
-The examination is made closely for the constructed variable month from Date.
-(@fig-month) is the boxplot showing the outliers, but there are not enough evidence to show that there are specific month that will have a higher lead concentration.
-Therefore, a similar filtering is applied to data larger than 2 ppm.
+### Graphs of Months vs Mean Lead Concentration
+
+With all data points, there are no obvious trends over time. The
+examination is made closely for the constructed variable month from
+Date. (@fig-month) is the boxplot showing the outliers, but there are
+not enough evidence to show that there are specific month that will have
+a higher lead concentration. Therefore, a similar filtering is applied
+to data larger than 2 ppm.
+
 ```{r}
@@
 
-Bar graph (@fig-meanmonth) shows the mean lead concentration for each month in the year.
-April and September has higher lead concentrations that is about 500% of the other months.
-However, this graph does not provide enough evidence to prove seasonality.
-It is necessary to examine whether April and September for every year will repeatedly have higher measured lead concentration.
-This will be further discussed in @sec-future.
+Bar graph (@fig-meanmonth) shows the mean lead concentration for each
+month in the year. April and September has higher lead concentrations
+that is about 500% of the other months. However, this graph does not
+provide enough evidence to prove seasonality. It is necessary to examine
+whether April and September for every year will repeatedly have higher
+measured lead concentration. This will be further discussed in
+@sec-future.
 
@@
 
-### graphs of years vs mean lead concentration
-Similar to months,variable called year was also constructed.
-However, both the boxplots and the mean values without data below 2 ppm does not show trend from year to year.
-This aligns with our conclusion in @sec-timeseries.
+### Graphs of Years vs Mean Lead Concentration
+
+Similar to months,variable called year was also constructed. However,
+both the boxplots and the mean values without data below 2 ppm does not
+show trend from year to year. This aligns with our conclusion in
+@sec-timeseries.
+
 ```{r}
@@
 ```
+
 ```{r}
@@
 
-## Weaknesses and next steps{#sec-future}
+## Weaknesses and Next Steps {#sec-future}
 
-A key weakness of the current dataset is that it originates from a variety of water sources and is not necessarily limited to direct drinking water.
-Since the water sample collection process was not controlled, it is uncertain whether all samples reflect tap water quality.
-Despite this, the guidelines used for analysis are intended for drinking water safety, which introduces a potential mismatch between the data and its intended use.
-The program is non-regulated, leaving the sampling process up to residents and without any standardization; as a result, the reliability and comparability of the data across different samples may be compromised.
+A key weakness of the current dataset is that it originates from a
+variety of water sources and is not necessarily limited to direct
+drinking water. Since the water sample collection process was not
+controlled, it is uncertain whether all samples reflect tap water
+quality. Despite this, the guidelines used for analysis are intended for
+drinking water safety, which introduces a potential mismatch between the
+data and our threshold maximum acceptable concentration. The program is
+non-regulated, leaving the sampling process up to residents and without
+any standardization. As a result, the reliability and comparability of
+the data across different samples may be varied.
 
-Considering these weaknesses, it is important to recognize the role of pre-treatment, such as water filtration systems, in affecting lead concentration.
-This necessitates the documentation of whether samples are taken before or after filtration, which could significantly influence the interpretation of lead levels in relation to health guidelines.
-To improve the reliability of future data collection, a probability-based adaptive sampling plan should be considered.
-This approach would ensure more representative sampling of tap water from various districts, providing a better understanding of the exposure to lead in drinking water.
-Additionally, further analysis on the seasonality of lead levels could reveal patterns influenced by environmental factors such as temperature, water flow rates, and seasonal water treatment processes.
-Identifying these trends is crucial for developing targeted interventions to reduce lead exposure throughout the year.
+It is important to recognize the role of pre-treatment, such as water
+filtration systems, in affecting lead concentration. The result of this
+paper might be altered if we can ensure every water sample was
+pre-treated to filter out the hard metals.
+
+To improve the reliability of future data analysis, a probability-based
+adaptive sampling plan should be considered. This approach would ensure
+more representative sampling of tap water from various districts,
+providing a better understanding of the exposure to lead in drinking
+water. There should also be probability based data analysis to better
+estimate correlations between lead concentrations and other variables,
+not limited to those mentioned in this paper.
+
+Additionally, further analysis on the seasonality of lead levels could
+verify whether there are patterns influenced by environmental factors
+such as temperature, water flow rates, and seasonal water treatment
+processes. Identifying these trends is crucial for developing targeted
+interventions to reduce lead exposure throughout for public health.
 
@@
 # Appendix {.unnumbered}
-The color used in graphs :
-blueishgrey <- rgb(96, 139, 182, maxColorValue = 255) is the color of lead in reality.
+
+The color used in graphs : blueishgrey \<- rgb(96, 139, 182,
+maxColorValue = 255) is the color of lead in reality.
+
+All outliers were represented by color = "red".
+
+This paper serves the purpose of telling story with data, there is only
+minimal analysis performed, for a deeper studies, consider use the
+processed data for training and analyzing.
 
 # Additional data details
+
+The data can be found in OpenDataToronto with
+<https://open.toronto.ca/dataset/non-regulated-lead-sample/>.
+
+The cleaning process includes removing NA, and modifying types of
+variables. Tests were performed to ensure that the data is suitable for
+analyze.
 
